---
title: "Splitting data into SOFT and HARD dasets - for NiN v 3 analyses - R Notebook"
output: html_notebook
---

#INPUTS
Whatever MAREANO dataset you are working with, it may be that there are too many hardbottom environments and taxa, and that these are confounding your M5 (soft bottom) analysis. If this is the case you may need to split your dataset into soft and hard datasets to analyse separately (as it is still useful to do an evaluation of M2 with the hard data)

This is done here using the sediment proportions, with a check of the taxon splits afterwards

PREVIOUSLY:

This is done using the taxonary to apply the split based on whether the taxa are obligate hard or soft bottom taxa. Any that could occur in either environment (e.g. those that are mobile or those where it is not so clear cut) are considered both and are retained in both datasets. This allows you to retain all samples, but only (roughly) the portions that are hard or soft separately.


## Libraries

```{r}
library(readxl)
library(tidyverse)
library(ggplot2)
library(gridExtra)
```


## load datasets to be split into hard and soft


```{r}
# PREFERABLY KEEP ALL LINES AND COMMENT OUT IF APPLYING TO NEW DATASET SO THAT THEY CAN BE TRACKED/REPEATED EASILY

# After SHELFoct1perR analysis, but now applied to all sample (Oslo) dataset
env_orig <- read.csv(file.path(dataPath,"inputsMS/Afot_env_sort_2023-08-04.csv")) %>% as.data.frame
otu_orig <- read.csv(file.path(dataPath,"inputsMS/Afot_otu_sort_2023-08-04.csv")) %>% as.data.frame
sedm_orig<- read.csv(file.path(dataPath,"inputsMS/sedimMeans.csv")) %>% as.data.frame

dim(env_orig)
dim(otu_orig)
dim(sedm_orig)
```
Sort out sedm_orig - pivot wider, allocate to hard sand/grav soft
```{r}
sedm_wide<-pivot_wider(sedm_orig,
                      id_cols="SampID",
                      names_from="vSubstrate",
                      values_from = "mean_percent")%>%
  select("SampID",sort(colnames(.))) #nice to have the cols alphabetical

names(sedm_wide)

env_orig<-left_join(env_orig, sedm_wide)
```

# prep helper columns

HI
```{r}

# HI sed proportion technique

vsoftCols<-c("vClay",   "vClay_spicules", "vMud", "vSandy_mud" )
vcoarseCols<-c("vSand",  "vSand_with_biomaterial", "vSand_with_shells", "vPebble")
vhardCols<-c("vBedrock", "vBoulder",  "vCobble", "vCompacted_sediment", "vKarbonatskorpe", "vDead_lophelia", "vCoral_rubble",  "vConsolidated_clay", "vClayOutcrop")




env_orig<-env_orig %>% 
  rowwise() %>%
  mutate(vHard=sum(c_across(all_of(vhardCols)),na.rm=T),
         vCoarse=sum(c_across(all_of(vcoarseCols)),na.rm=T),
         vSoft=sum(c_across(all_of(vsoftCols)),na.rm=T))

plot(y=env_orig$vHard, x=env_orig$VL)
plot(env_orig$vSoft)
plot(env_orig$vCoarse)




```


NGU cats
```{r}
# NGU sed class technique

softNGUCats<-c("10", "15", "20", "21", "30", "40", "50", "60", "100", "115")
coarseNGUCats<-c("70", "80","90", "95", "100", "105", "120", "130", "140", "150", "160", "170")
hardNGUCats<-c("1", "5", "174", "175", "180", "185", "190", "205", "206", "210", "215", "300", "500")

plot(as.factor(env_orig$sedclass))
```

NGU Props

```{r}

env_orig<-env_orig %>% 
  mutate (HardNGU=sum(c_across(all_of(c("cobB", "rock"))),na.rm=T),
          CoarseNGU=sum(c_across(all_of(c("gravel", "sand"))),na.rm=T),
          SoftNGU=mud)

plot(env_orig$HardNGU)
plot(env_orig$SoftNGU)
plot(env_orig$CoarseNGU)
```



#Split based on sample sediment proportions

We are trying this three different ways to see what the difference in results is
```{r}

# HI sed proportion technique

env_orig<-env_orig %>% 
  mutate (sedPropCatHI=if_else(vHard>vCoarse & vHard>vSoft, "Hard", if_else(vSoft>vCoarse & vSoft>vHard, "Soft",if_else(vCoarse>vSoft & vCoarse>vHard,"Coarse", "na"))))

print("HI sed proportions - sample predominantly:")
table(env_orig$sedPropCatHI)


# NGU Sed class technique

env_orig<-env_orig %>%
  mutate (sedclassCatNGU=if_else(sedclass %in% hardNGUCats, "Hard", if_else(sedclass %in% softNGUCats, "Soft",if_else(sedclass %in% coarseNGUCats,"Coarse","na"))))


# env_orig<-env_orig %>%
#   mutate (sedclassCatNGU=if_else((sedclass>173)|(sedclass>0 & sedclass<6), "Hard", if_else(sedclass>5 & sedclass<94, "Soft","Coarse")))

# env_orig<-env_orig %>% 
#   mutate (sedclassCatNGU=if_else(HardNGUClass>CoarseNGUClass & HardNGUClass>SoftNGUClass, "Hard", if_else(SoftNGUClass>CoarseNGUClass & SoftNGUClass>HardNGUClass, "Soft","Coarse")))

print("NGU sed categories - sample predominantly:")
table(env_orig$sedclassCatNGU)

# NGU sed proportion technique


env_orig<-env_orig %>% 
  mutate (sedPropCatNGU=if_else(HardNGU>CoarseNGU & HardNGU>SoftNGU, "Hard", if_else(SoftNGU>CoarseNGU & SoftNGU>HardNGU, "Soft",if_else(CoarseNGU>SoftNGU & CoarseNGU>HardNGU,"Coarse", "na"))))

print("NGU sed proportions - sample predominantly:")
table(env_orig$sedPropCatNGU)


```

Plot to compare where diff techniques see as hard/coarse/soft
```{r}
#HI technique
p_HIvsed <- ggplot(data = env_orig,
                     aes(x = x_coord,
                         y = y_coord)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("HI sediment proportions classified as Hard/Coarse/Soft",
          subtitle = "First run") +
  geom_point(aes(colour = factor(sedPropCatHI))) +
  # geom_vline(xintercept = 0,
  #            linetype = 2,
  #            colour = "lightgray") +
  # geom_hline(yintercept = 0,
  #            linetype = 2,
  #            colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

p_HIvsed
```
```{r}
#NGU class technique
p_NGUclsed <- ggplot(data = env_orig,
                     aes(x = x_coord,
                         y = y_coord)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("NGU sediment classes classified as Hard/Coarse/Soft",
          subtitle = "First run") +
  geom_point(aes(colour = factor(sedclassCatNGU))) +
  # geom_vline(xintercept = 0,
  #            linetype = 2,
  #            colour = "lightgray") +
  # geom_hline(yintercept = 0,
  #            linetype = 2,
  #            colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

p_NGUclsed
```



```{r}
#NGU proportions technique
p_NGUpropsed <- ggplot(data = env_orig,
                     aes(x = x_coord,
                         y = y_coord)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("NGU sediment proportions classified as Hard/Coarse/Soft",
          subtitle = "First run") +
  geom_point(aes(colour = factor(sedPropCatNGU))) +
  # geom_vline(xintercept = 0,
  #            linetype = 2,
  #            colour = "lightgray") +
  # geom_hline(yintercept = 0,
  #            linetype = 2,
  #            colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

p_NGUpropsed
```


```{r}
grid.arrange(p_NGUpropsed,  p_NGUclsed, p_HIvsed, ncol=3)
```
Add concept of mixed to sediment categorisations



```{r}

# HI sed proportion technique

env_orig<-env_orig %>% 
  mutate (sedPropCatHImix=if_else(vHard>0.5, "Hard", if_else(vSoft>0.5, "Soft",if_else(vCoarse>0.5,"Coarse", "Mixed"))))

print("HI sed proportions inc mix - sample predominantly:")
table(env_orig$sedPropCatHImix)


# NGU Sed class technique


# softNGUCats<-c("10", "15", "20", "21", "30", "40", "50", "60", "100", "115")
# coarseNGUCats<-c("70", "80","90", "95", "100", "105", "120", "130", "140", "150", "160", "170")
# hardNGUCats<-c("1", "5", "174", "175", "180", "185", "190", "205", "206", "210", "215", "300", "500")

#assume that if used first in if_else list will be labelled mix even if repeated in other category list (above)
mixNGUCats<-c("1","185","190","205","206","210", "215","500")





env_orig<-env_orig %>%
  mutate (sedclassCatNGUmix=if_else(sedclass %in% mixNGUCats, "Mixed",if_else(sedclass %in% hardNGUCats, "Hard", if_else(sedclass %in% softNGUCats, "Soft",if_else(sedclass %in% coarseNGUCats,"Coarse","na")))))


# env_orig<-env_orig %>%
#   mutate (sedclassCatNGU=if_else((sedclass>173)|(sedclass>0 & sedclass<6), "Hard", if_else(sedclass>5 & sedclass<94, "Soft","Coarse")))

# env_orig<-env_orig %>% 
#   mutate (sedclassCatNGU=if_else(HardNGUClass>CoarseNGUClass & HardNGUClass>SoftNGUClass, "Hard", if_else(SoftNGUClass>CoarseNGUClass & SoftNGUClass>HardNGUClass, "Soft","Coarse")))

print("NGU sed categories inc mix - sample predominantly:")
table(env_orig$sedclassCatNGUmix)














# NEED TO FIX THE BELOW WHICH IS DOING SOMETHING WEIRD...


# [1] "HI sed proportions inc mix - sample predominantly:"
# 
# Coarse   Hard  Mixed   Soft 
#    122   3759     45    720 
# [1] "NGU sed categories inc mix - sample predominantly:"
# 
# Coarse   Hard  Mixed   Soft     na 
#   2041    237    587   1770     11 
# [1] "NGU sed inc mix proportions - sample predominantly:"
# 
# Hard Soft 
#  939 3707 



# NGU sed proportion technique


env_orig<-env_orig %>% 
  mutate (sedPropCatNGUmix=if_else(HardNGU>0.5, "Hard", if_else(CoarseNGU>0.5,"Coarse", if_else(SoftNGU>0.5, "Soft", "Mixed"))))

print("NGU sed inc mix proportions - sample predominantly:")
table(env_orig$sedPropCatNGUmix)


```


# TAXONARY PREP
## load taxonary and make hard/soft lookup tables
NB Up-to-date live version of taxonary available at dataPath_2, but snapshot version from 20.10.22 is available in dataPath (includes hardSoft checked by bex & gjertrud)

was previously using this as the primary split method, but cannot assign without knowing diversity ratios. Suspecting 3:1 hard diverity:soft diversity. 

```{r}
# taxonary <- read_xlsx(file.path(dataPath_2, "Taxonary.xlsx"), sheet=1) %>%
#   mutate(clean_taxonomy = gsub(" ","_",`Reference List`)) %>%#ensure names match format in otu_orig
#   mutate(clean_taxonomy = gsub("\\;",".",clean_taxonomy, fixed=FALSE))

taxonary <- read_xlsx(file.path(dataPath, "inputsMS/Taxonary_oct2022.xlsx"), sheet=1)%>%
 mutate(clean_taxonomy = gsub(" ","_",`Reference List`)) %>%#ensure names match format in otu_orig
 mutate(clean_taxonomy = gsub("\\;",".",clean_taxonomy, fixed=FALSE)) 


taxonary <- taxonary  %>%
  mutate(across('clean_taxonomy', str_replace,'Hexactinellida_urn-shaped', 'Hexactinellida_urn.shaped'))


# AMEND SOMTIME

# Warning: There was 1 warning in `mutate()`.
# ℹ In argument: `across(...)`.
# Caused by warning:
# ! The `...` argument of `across()` is deprecated as of dplyr 1.1.0.
# Supply arguments directly to `.fns` through an anonymous function instead.
# 
#   # Previously
#   across(a:b, mean, na.rm = TRUE)
# 
#   # Now
#   across(a:b, \(x) mean(x, na.rm = TRUE))
```

# SPLIT DATASET INTO SOFT/HARD

## Format datasets for fitering spp
```{r}
otu_orig[otu_orig==0]<-NA
otu_long<-otu_orig %>% pivot_longer(cols=!c("X","SampID"),
                                    names_to="Taxon",
                                    values_to="Density",
                                    values_drop_na=T)
unique(otu_long$Taxon) #list of species to see what to expect
```

## Split taxonary into hard/soft lookup tables
Will retain only "soft" and "both" categories in the soft dataset
Will retain only "hard" and "both" categories in the hard dataset

N/A applies to pelagic species and objects to ensure they are also removed
```{r}
# softsppincNA<-subset(taxonary,HardSoft_Affiliated!="Hard")
# softspp<-subset(softsppincNA,HardSoft_Affiliated!="N/A")
# print("Soft includes..")
# unique(softspp$HardSoft_Affiliated)
# table(softspp$HardSoft_Affiliated)
# 
# hardsppincNA<-subset(taxonary,HardSoft_Affiliated!="Soft")
# hardspp<-subset(hardsppincNA,HardSoft_Affiliated!="N/A")
# print("Hard includes..")
# unique(hardspp$HardSoft_Affiliated)
# table(hardspp$HardSoft_Affiliated)


softsppincNA<-subset(taxonary,HardSoft_Affiliated=="Soft")
softspp<-subset(softsppincNA,HardSoft_Affiliated!="N/A")
print("Soft includes..")
unique(softspp$HardSoft_Affiliated)
table(softspp$HardSoft_Affiliated)

hardsppincNA<-subset(taxonary,HardSoft_Affiliated=="Hard")
hardspp<-subset(hardsppincNA,HardSoft_Affiliated!="N/A")
print("Hard includes..")
unique(hardspp$HardSoft_Affiliated)
table(hardspp$HardSoft_Affiliated)
```

## filter otu_long by hard/soft taxa

```{r}
otu_long_hard<-otu_long %>%filter(Taxon %in% hardspp$clean_taxonomy)
print("HARD--------------")
n_distinct(otu_long_hard$Taxon)
unique(otu_long_hard$Taxon) #list of hard species to see what to expect
otu_long_soft<-otu_long %>%filter(Taxon %in% softspp$clean_taxonomy)
print("SOFT--------------")
n_distinct(otu_long_soft$Taxon)
unique(otu_long_soft$Taxon) #list of soft species to see what to expect (number restarts from 1)
```
## Check for skipped samples/species

```{r}
otu_long_skipped<-otu_long %>%
  filter(!Taxon %in% hardspp$clean_taxonomy) %>% 
  filter(!Taxon %in% softspp$clean_taxonomy)

otu_long_skipped
```


## pivot wider again
```{r}
otu_hard<-pivot_wider(otu_long_hard,
                      id_cols="SampID",
                      names_from="Taxon",
                      values_from = "Density",
                      values_fill = 0)%>%
  select("SampID",sort(colnames(.))) #nice to have the cols alphabetical

otu_soft<-pivot_wider(otu_long_soft,
                      id_cols="SampID",
                      names_from="Taxon",
                      values_from = "Density",
                      values_fill = 0)%>%
  select("SampID",sort(colnames(.))) #nice to have the cols alphabetical
```


## Decide if sample predominantly hard or soft (new step)
```{r}
#Taxon technique
count_hard<-otu_hard%>% mutate(countHard=(rowSums(.>0))) %>% select(c(SampID, countHard))
count_soft<-otu_soft%>% mutate(countSoft=(rowSums(.>0))) %>% select(c(SampID, countSoft))

env_orig<- left_join(env_orig,count_hard, by="SampID")
env_orig<- left_join(env_orig,count_soft, by="SampID")

env_orig<- env_orig %>% 
  mutate (countHard=coalesce(countHard,0),
          countSoft=coalesce(countSoft,0),
    PredomTaxaHardSoft=if_else(countHard>(3*countSoft),"Hard","Soft"))

table(env_orig$PredomTaxaHardSoft)


#Create Hard, Soft and Mix categories:
# Hard if > 80% hard
# Soft if <20% soft
# Mix if 20-80% hard

env_orig<-env_orig %>% 
  mutate (propHard=countHard/(countHard+countSoft),
          HardSoftMix=if_else(propHard>0.90, "Hard", if_else(propHard<0.50, "Soft","Mix")))

table(env_orig$HardSoftMix)

leftovers<-subset(env_orig, is.na(PredomTaxaHardSoft))
leftovers
```



compare plot of hard soft mix with sediment plots

```{r}
#NGU proportions technique
p_taxaPropHSMix <- ggplot(data = env_orig,
                     aes(x = x_coord,
                         y = y_coord)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("NGU sediment proportions classified as Hard/Coarse/Soft",
          subtitle = "First run") +
  geom_point(aes(colour = factor(HardSoftMix))) +
  # geom_vline(xintercept = 0,
  #            linetype = 2,
  #            colour = "lightgray") +
  # geom_hline(yintercept = 0,
  #            linetype = 2,
  #            colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

p_taxaPropHSMix
```







## Match hard/soft otu sample data to env sample data and check we labelled everything...
```{r}
env_hard<-env_orig %>% filter(PredomHardSoft=="Hard")
env_soft<-env_orig %>% filter(PredomHardSoft=="Soft")

otu_hard<-otu_hard %>% filter(SampID%in%env_hard$SampID)
otu_soft<-otu_soft %>% filter(SampID%in%env_soft$SampID)

print("Number of samples in hard dataset...")
nrow(env_hard)
print("Number of samples in soft dataset...")
nrow(env_soft)

print("Total Number of samples in hard/soft datasets...")
sum(nrow(env_hard)+nrow(env_soft))
print("Total Number of samples in env_orig...")
length(unique(env_orig$SampID))

```


```{r}
# env_hard<-env_orig%>% filter(SampID%in%otu_hard$SampID) 
# #check lengths match
# length(unique(env_hard$SampID)) # samp in env dataset
# length(unique(otu_hard$SampID)) # samp in otu dataset
# 
# env_soft<-env_orig%>% filter(SampID%in%otu_soft$SampID) 
# #check lengths match 
# length(unique(env_soft$SampID)) # samp in env dataset
# length(unique(otu_soft$SampID)) # samp in otu dataset

```
# Write HARD and SOFT datasets
```{r}
# BE SURE TO EDIT FILENAMES

# HARD DATSET
write.csv(otu_hard, file=file.path(dataPath,(paste0("inputsMS/afot_hard_otu_", Sys.Date(),".csv"))))
write.csv(env_hard, file=file.path(dataPath,(paste0("inputsMS/afot_hard_env_", Sys.Date(),".csv"))))
# SOFT DATASNAW
write.csv(otu_soft, file=file.path(dataPath,(paste0("inputsMS/afot_soft_otu_", Sys.Date(),".csv"))))
write.csv(env_soft, file=file.path(dataPath,(paste0("inputsMS/afot_soft_env_", Sys.Date(),".csv"))))
```




