---
title: "VISUALISE SPLITS - MAREANO, Deep Sea NiN v 3.0 R Notebook"
output: html_notebook
---

After deciding all splits, this script uses the rules applied to make the splits in order to generate a single dataset with all splits labelled.

# Inputs
 Need the all samples dataset and the rules needed to apply each split.
 
## Libraries
```{r}
library(tidyverse)
library(oce)
library(RColorBrewer)
library(plotly)
```

## Load data
 
```{r}
#NB 2022-09-29 was the topcopy (all sample/Oslo) dataset being used before the latest species cleaning
#NB 2022-10-19 latest (all sample/Oslo) species cleaning applied so smaller spp dataset

# USE DIFFERENT SCRIPT FOR 1perR station dataset

env_orig <- read.csv(file.path(dataPath,"inputs/envDataOslo_2022-10-19.csv")) %>% as.data.frame
otu_orig <- read.csv(file.path(dataPath,"inputs/widedataOslo_2022-10-19.csv")) %>% as.data.frame



#left join and split to ensure ordered identically

#joinedDat<-otu_orig%>% dplyr::left_join(.,env_orig, by="SampID")
joinedDat<- left_join(otu_orig,env_orig, by=c("SampID"="SampID2"))
```

## Add variables for explorations


### sea water density
```{r}
joinedDat$swDensRob_avs<-swRho(salinity=joinedDat$Smean_Robinson,
                         temperature=joinedDat$Tmean_Robinson,
                         pressure=(joinedDat$bathy*-1),
                         eos="unesco")
joinedDat$bathyPositive<-joinedDat$bathy*-1
```

### MLD-bathy vars
```{r}
joinedDat<-joinedDat %>% 
  mutate ("MLDmean_bathy"=MLDmean_Robinson-(bathy*-1),
          "MLDmin_bathy"=MLDmin_Robinson-(bathy*-1),
          "MLDmax_bathy"=MLDmax_Robinson-(bathy*-1))

joinedDat$MLDmean_bathy<-cut(joinedDat$MLDmean_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))
joinedDat$MLDmin_bathy<-cut(joinedDat$MLDmin_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))
joinedDat$MLDmax_bathy<-cut(joinedDat$MLDmax_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))
```


### EUNIS, seddan, sedmil

You can look up EUNIS column meanings here - p14 on: https://archimer.ifremer.fr/doc/00723/83528/88577.pdf
Sediment miljø (erosion/deposition stuff) https://www.ngu.no/upload/Kartkatalog/Presentasjonsregler_Marin_Sedimentasjonsmiljo.pdf
Sediment dannelse (genesis) https://www.ngu.no/upload/Kartkatalog/Presentasjonsregler_Marin_SedimentDannelse.pdf
```{r}
envExtra<-read.csv(file.path(dataPath,"inputs/samples_env_plus.csv")) %>% as.data.frame
envExtra<-envExtra %>% select(c("SampID2", 
                                "Energy",#EUNIS #TEXT
                                "Biozone",#EUNIS #TEXT
                                "Substrate",#EUNIS #TEXT
                                "EUNIScomb",#EUNIS #TEXT
                                "EUNIScombD",#EUNIS #TEXT
                                "Allcomb",#EUNIS #TEXT
                                "AllcombD",#EUNIS #TEXT
                                "MSFD_BBHT",#EUNIS #TEXT
                                "EUNIS2019C",#EUNIS #TEXT
                                "EUNIS2019D",#EUNIS #TEXT
                                "All2019D",#EUNIS #TEXT
                                "All2019DL2",#EUNIS #TEXT
                                "sedmil", #erosion/deposition #CATEGORICAL
                                "seddan", #sed genesis #CATEGORICAL 
                                "sw_spd_max", #sandwave model max current speed
                                "sw_spd_mean", #sandwave model mean current speed
                                "TOC0.10cm_median" #Total Organic Carbon (Markus Diesing preliminary model 10/2022 top 10cm prediction)
))


#save envEtra for later, but add TOC now for use in correlations
envTOC<-envExtra %>% select(c("SampID2", "TOC0.10cm_median"))

joinedDat<-left_join(joinedDat,envTOC, by=c("SampID"="SampID2"))
```



### video sediment fractions
The list of video fractions needs consolidating to be backwards compatible with earlier records to avoid NAs



```{r}
sedVid_long<-read.csv(file.path(dataPath,"inputs/sedimMeans.csv")) %>% as.data.frame
sedVid<-sedVid_long %>% pivot_wider(id_cols=SampID,
                                    names_from = vSubstrate,
                                    values_from=mean_percent,
                                    values_fill=NA)

joinedDat<-left_join(joinedDat,sedVid, by=c("SampID"))



```






# Label splits

The following rules need to be applied:

Barents Sea - Ice:

* <1030 density, >305 dissox, >0 max ice cover

Barents Sea - No ice (inc WBE):
* <1030 density, >dissox 300.5 dissox, 0 max ice cover (NB WBE was dissox 300.5 - 305)


Atlantic (inc WBW):

* <1030 density, <300.5 dissox (NB WBW was 294-300.5 dissox)

Deep shelf break (excl Aktivneset):

* >1030 density, >2.5*C (retained (x_coord>30000)&(x_coord<600000))


Deep cold:

* >1030 density, <2.5*C

***NB SOME INDIVIDUALLY IDENTIFIED OUTLIERS MAY NOT BE LABELLED YET***


```{r}
joinedDat <- joinedDat %>% 
  mutate("split"=case_when(
    bathyPositive>100 & swDensRob_avs<1030 & BO22_dissoxmean_bdmean>300.5 & BO22_icecoverltmax_ss>0~"Barents_Ice", #technically >305 but wanted to capture the ice cover that occurred at 305
    bathyPositive>100 & swDensRob_avs<1030 & BO22_dissoxmean_bdmean>300.5 & BO22_icecoverltmax_ss==0~"Barents_noICE",
    bathyPositive>100 & swDensRob_avs<1030 & BO22_dissoxmean_bdmean<=300.5 & cruise_no!=2006612~"Atlantic",
    bathyPositive>100 & swDensRob_avs>=1030 & Tmean_Robinson>2.5 & (x_coord>30000 &x_coord<600000)~"Deep_Warm",
    bathyPositive>100 & swDensRob_avs>=1030 & Tmean_Robinson<=2.5~"Deep_Cold",
    bathyPositive<=100~"OUTLIERS<100m",
   TRUE~"OUTLIERS-other"
         
))

summary(as.factor(joinedDat$split))
```


#Map the splits

### a colour palette for discrete plots (see later in script)
```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# discrete colur pallete for up to 74 colours/categories
n <- 74
qual_col_pals <-brewer.pal.info[brewer.pal.info$category == 'qual',]

col_vector <-unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

pie(rep(1,n), col=sample(col_vector, n))
```

#### X v Y with gnmds ax 1 as HC categories
```{r}
map_spl <- ggplot(data = joinedDat,
              aes(x = x_coord,
                  y = y_coord)) +
  theme_classic() +
  geom_point(aes(colour = joinedDat$split), #change at will - categorical vars only
             size = 1) +
  scale_colour_manual(values=col_vector) +# non-ordered colourblind pallette
   scale_colour_manual(values=c("Atlantic"="gold",
                               "Barents_Ice"="springgreen",
                               "Barents_noICE"="green4",
                               "Deep_Cold"="navy",
                               "Deep_Warm"="royalblue1",
                               "OUTLIERS<100m"="gray70",
                               "OUTLIERS-other"="gray40"
                               )) +
 # scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by Split") #amend

#cat_xy_ax1 #plot
ggplotly(map_spl) #interactive plot


```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/Splits_plot.png"), #amend
       device = "png",
       dpi=300 )

```
#### X v Y - Lophelia
```{r}
xy_t<- ggplot(data = joinedDat,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Lophelia_pertusa),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(0.0001, max(joinedDat$Lophelia_pertusa)), #match variable acting as colour
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by Lophelia pertusa") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_WB_W_LOPH__X_v_Y_loph.png"), #amend
       device = "png",
       dpi=300 )
```



***temporary***

## explore video sediments

```{r}
crvl<-joinedDat %>% select(SampID,cruise_no,VL)

sedVid_long<-left_join(sedVid_long, crvl)


table(sedVid_long$cruise_no,sedVid_long$Substrate)


```














































 
 