---
title: "Splitting data in 1sample per R station dataset - for NiN v 3 analyses - R Notebook"
output: html_notebook
---

#INPUTS
Assumes you hve already run at least the scripts:
- workflow_deepseaNiN_Start
- Workflow_deepseaNiN_Start_splitBiogeog

(Else further subdivisions of the same.) 

Requires the main analysis environment to be loaded.

# Selecting geographic subsets
This is to explore thresholds/relationships locally. E.g. To look for threshold values/ local relationships before deciding on splits. You need to already have a theory that you want to test in order to select the geographic regions to look at.



## OPTIONAL: load new dataset and apply splits SKIP IF ALREADY HAVE "joinedDat"

CURRENTLY DESIGEND TO JUST SUPPLY THE SHELF DATASET IN 1 SAMP PER R STATION

```{r}
#env_orig <- read.csv(file.path(dataPath,"inputs/envDataOslo_2022-09-29.csv")) %>% as.data.frame
#otu_orig <- read.csv(file.path(dataPath,"inputs/widedataOslo_2022-09-29.csv")) %>% as.data.frame


#NB 2022-09-13 was the topcopy (1perR) dataset being used before the latest species cleaning
#2022-10-19 latest (1perR) species cleaning applied
# USE DIFFERENT SCRIPT FOR "all samples" (Oslo) station dataset

env_orig <- read.csv(file.path(dataPath,"inputs/envData_2022-09-13.csv")) %>% as.data.frame
otu_orig <- read.csv(file.path(dataPath,"inputs/widedata_2022-09-13.csv")) %>% as.data.frame

#left join and split to ensure ordered identically

#joinedDat<-otu_orig%>% dplyr::left_join(.,env_orig, by="SampID")
joinedDat<- left_join(otu_orig,env_orig, by=c("SampID"="SampID2"))

```




## MAIN SPLITS

### Density threshold splits

density split 1031.8 threshold

```{r}
joinedDat$swDensRob_avs<-swRho(salinity=joinedDat$Smean_Robinson,
                         temperature=joinedDat$Tmean_Robinson,
                         pressure=(joinedDat$bathy*-1),
                         eos="unesco")

loDensShallowJoinedDat<-subset(joinedDat,swDensRob_avs<1031.1)
plot(loDensShallowJoinedDat$X.y, loDensShallowJoinedDat$Y)
LDensShlw_otu_sort<-loDensShallowJoinedDat %>% select(c(2:351))
LDensShlw_env_sort<-loDensShallowJoinedDat %>% select(c(2,357:491))                                


# hiDensDeepJoinedDat<-subset(joinedDat,swDensRob_avs>1031.1)
# plot(hiDensDeepJoinedDat$X.y, hiDensDeepJoinedDat$Y)
# HDensDeep_otu_sort<-hiDensDeepJoinedDat %>% select(c(2:351))
# HDensDeep_env_sort<-hiDensDeepJoinedDat %>% select(c(2,357:491))  
```


##### SAVE DATASETS
```{r}
# write.csv(HDensDeep_otu_sort, file=file.path(dataPath,(paste0("inputs/HDensDeep_otu_sort_", Sys.Date(),".csv"))))
# write.csv(HDensDeep_env_sort, file=file.path(dataPath,(paste0("inputs/HDensDeep_env_sort_", Sys.Date(),".csv"))))

# write.csv(LDensShlw_otu_sort, file=file.path(dataPath,(paste0("inputs/LDensShlw_otu_sort_", Sys.Date(),".csv"))))
# write.csv(LDensShlw_env_sort, file=file.path(dataPath,(paste0("inputs/LDensShlw_env_sort_", Sys.Date(),".csv"))))
```




### REMOVE SPITSBERGENBANKEN OUTLIERS FROM LowDens Dataset

c("1805_01","1801_00","1802_02", "1799_01", "1803_00", "1798_02","1800_00","1995_00")

```{r}
#SB<-c("1805_01","1801_00","1802_02", "1799_01", "1803_00", "1798_02","1800_00","1995_00")

plot(loDensShallowJoinedDat$X.y, loDensShallowJoinedDat$Y)

#loDensNoSBJoinedDat <-loDensShallowJoinedDat %>% filter(!(SampID %in% c("1805_01",
                                                                        # "1801_00",
                                                                        # "1802_02", 
                                                                        # "1799_01", 
                                                                        # "1803_00", 
                                                                        # "1798_02",
                                                                        # "1800_00",
                                                                        # "1995_00")))
loDensShallowJoinedDat$bathyPositive<-(loDensShallowJoinedDat$bathy*-1)
loDensNoSBJoinedDat <-subset(loDensShallowJoinedDat, bathyPositive>100)
loDensNoSBJoinedDat<-loDensNoSBJoinedDat%>%select(-c("bathyPositive"))



plot(loDensNoSBJoinedDat$X.y, loDensNoSBJoinedDat$Y)


LDensNoSB_otu_sort<-loDensNoSBJoinedDat %>% select(c(2:351))
LDensNoSB_env_sort<-loDensNoSBJoinedDat %>% select(c(2,357:491))  

```
##### SAVE DATASETS
```{r}
# write.csv(LDensNoSB_otu_sort, file=file.path(dataPath,(paste0("inputs/LDensNoSB_otu_sort_", Sys.Date(),".csv"))))
# write.csv(LDensNoSB_env_sort, file=file.path(dataPath,(paste0("inputs/LDensNoSB_env_sort_", Sys.Date(),".csv"))))
```


### Split LoDensNoSB by dissox

```{r}
#SB<-c("1805_01","1801_00","1802_02", "1799_01", "1803_00", "1798_02","1800_00","1995_00")

plot(loDensNoSBJoinedDat$X.y, loDensNoSBJoinedDat$Y)


# loDensNoSBhiO2JoinedDat<-subset(loDensNoSBJoinedDat,BO22_dissoxmean_bdmean>305)
# plot(loDensNoSBhiO2JoinedDat$X.y, loDensNoSBhiO2JoinedDat$Y)
# LDnoSBhi02_otu_sort<-loDensNoSBhiO2JoinedDat %>% select(c(2:351))
# LDnoSBhi02_env_sort<-loDensNoSBhiO2JoinedDat %>% select(c(2,357:491))  


loDensNoSBloO2JoinedDat<-subset(loDensNoSBJoinedDat,BO22_dissoxmean_bdmean<305)
plot(loDensNoSBloO2JoinedDat$X.y, loDensNoSBloO2JoinedDat$Y)
LDnoSBlo02_otu_sort<-loDensNoSBloO2JoinedDat %>% select(c(2:351))
LDnoSBlo02_env_sort<-loDensNoSBloO2JoinedDat %>% select(c(2,357:491))    

```
##### SAVE DATASETS
```{r}
write.csv(LDnoSBlo02_otu_sort, file=file.path(dataPath,(paste0("inputs/LDnoSBlo02_otu_sort_", Sys.Date(),".csv"))))
write.csv(LDnoSBlo02_env_sort, file=file.path(dataPath,(paste0("inputs/LDnoSBlo02_env_sort_", Sys.Date(),".csv"))))

# write.csv(LDnoSBhi02_otu_sort, file=file.path(dataPath,(paste0("inputs/LDnoSBhi02_otu_sort_", Sys.Date(),".csv"))))
# write.csv(LDnoSBhi02_env_sort, file=file.path(dataPath,(paste0("inputs/LDnoSBhi02_env_sort_", Sys.Date(),".csv"))))
```



