---
title: "MAREANO - NiN: First run"
authors: Thijs van Son, Rune Halvorsen, Rebecca Ross, Genoveva Gonzales-Mirelis, Margaret Dolan
date: "Last Rendered on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 2
    toc_float: yes
    fig_width: 7
    fig_height: 7
always_allow_html: true 
---

## Packages
```{r}
library(dplyr)
library(ggcorrplot)
library(ggpubr)
library(patchwork)
# library("here")
# library("bookdown")
# library("downloadthis")
library(vegan)
library(plyr)
library(e1071)
library(tidyverse)
library(viridis)
library(GGally)
library(ggrepel)
library(ggordiplots)
library(readr)
library(RColorBrewer)
library(oce) 


source("gg_ordisurf_viridis.R")
```

## Load R objects from previous runs if needed

Only do this if you have already run this script and would like to avoid rerunning the longer steps
```{r}
geodist_pa<-readRDS(file.path(dataPath,"inputs/geodist_pa.rds"))
geodist_r6<-readRDS(file.path(dataPath,"inputs/geodist_r6.rds"))

mds_pa<-readRDS(file.path(dataPath,"inputs/mds_pa.rds"))
mds_r6<-readRDS(file.path(dataPath,"inputs/mds_r6.rds"))
mds_r6_1000<-readRDS(file.path(dataPath,"inputs/mds_r6_1000rep.rds"))

ep<-0.8 #epsilon (to avoid having to find and run just this line from code blocks where the mds objects were made)
```


## Data
```{r}

env_orig <- read.csv(file.path(dataPath,"inputs/envData_2022-09-13.csv")) %>% as.data.frame
otu_orig <- read.csv(file.path(dataPath,"inputs/widedata_2022-09-13.csv")) %>% as.data.frame

#left join and split to ensure ordered identically

#joinedDat<-otu_orig%>% dplyr::left_join(.,env_orig, by="SampID")
joinedDat<- left_join(otu_orig,env_orig, by=c("SampID"="SampID2"))

otu_sort<-joinedDat %>% select(c(2:351))
env_sort<-joinedDat %>% select(c(2,352:491))

# Sorting and ordering by SampID
# env_orig <- env_orig[order(env_orig$SampID), ]
# otu_orig <- otu_orig[order(otu_orig$SampID), ]

table(env_sort$SampID2 == otu_sort$SampID)


#must be same length and equal to unique sample number length - the ordered lists are assumed to be directly relatable on a row by row basis

dim(env_sort)
dim(otu_sort)
```

#### remove any variables with not enough coverage
```{r}

env_sort <- env_sort %>% select(-c("BO22_lightbotmean_bdmean",
                               "BO22_lightbotltmax_bdmean",
                               "BO22_lightbotltmin_bdmean",
                               "BO22_lightbotrange_bdmean"))
```


#### Plot map to check location of samples
```{r}
env_sort_locations<- ggplot(data = env_sort,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Location of samples in sorted env file")

env_sort_locations
```


### Data cleaning
```{r}
## Removing NAs
otuCompl <- otu_sort[complete.cases(env_sort[, -c(1:20)]), ]
envCompl <- env_sort[complete.cases(env_sort[, -c(1:20)]), ]

## Removing observations with less than 4 OTUs
sel <- rowSums(otuCompl[, -c(1:2)]) >= 4
otuSel <- otuCompl[sel, ]
envSel <- envCompl[sel, ]


## Removing phosphate
#envSel <- envSel %>% select(-phosphate_mean.tif)


dim(otuSel); dim(envSel)
```
#### Show what samples are left after complete cases and >4 OTUs filters
```{r}
env_sel_locations<- ggplot(data = envSel,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Location of samples in sorted env file")

env_sel_locations
```

#### Show what got removed in complete cases filter
```{r}
envCompl_ccrem<-env_sort%>%filter(!SampID%in%envCompl$SampID)

envCompl_ccrem_locations<- ggplot(data = envCompl_ccrem,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Location of removed complete case samples resulting in envSel file")

envCompl_ccrem_locations
```
#### Show what got removed in <4 OTUs filter


```{r}
inv.sel <- rowSums(otuCompl[, -c(1:2)]) < 4
env.invSel <- envCompl[inv.sel, ]

env.invSel_locations<- ggplot(data = env.invSel,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Location of removed samples with <4 OTUs resulting in envSel file")

env.invSel_locations

```

### [Optional] Data thinning
```{r, eval=FALSE}
# otu_red <- otu1[1:500, ]
# otu <- otu_red
# 
# env_red <- env[1:500, ]
# env <- env_red
```

### Whole dataset for first run
```{r}
otu<-otuSel[,-c(1)]
env<-envSel[, -c(2:22)]

table(is.na(otu))

table(is.na(env))

#str(otuSel)
#str(envSel)
```


### Splitting in subsets
```{r}
# ## Class 1
# otu1 <- subset(otuSel, envSel$SplitRev == 1)
# env1 <- envSel %>% filter(SplitRev == 1)
# 
# ## Class 2
# otu2 <- subset(otuSel, envSel$SplitRev == 2)
# env2 <- envSel %>% filter(SplitRev == 2)
# 
# ## Class 4
# otu3 <- subset(otuSel, envSel$SplitRev == 3)
# env3 <- envSel %>% filter(SplitRev == 3)
# 
# ## Class 4
# otu4 <- subset(otuSel, envSel$SplitRev == 4)
# env4 <- envSel %>% filter(SplitRev == 4)
# 
# ## Class 6
# otu6 <- subset(otuSel, envSel$SplitRev == 6)
# env6 <- envSel %>% filter(SplitRev == 6)
# 
# ## Class 7
# otu7 <- subset(otuSel, envSel$SplitRev == 7)
# env7 <- envSel %>% filter(SplitRev == 7)
# 
# ## Class 8
# otu8 <- subset(otuSel, envSel$SplitRev == 8)
# env8 <- envSel %>% filter(SplitRev == 8)
```

## Selecting subset
```{r}
# ## Selecting 
# otu <- otu2
# env <- env2

```

## format data correctly
```{r}
#env$coords.x1<-as.numeric(env$coords.x1)
#env$coords.x2<-as.numeric(env$coords.x2)


```


## Abundance weighting
Make function
You might have to drop variables that have been imported as character
```{r}

otu_pa <- decostand(x = otu[, -c(1)],
                    method = "pa")

# y = ax^w         # power transformation formula
dt <- otu[, -c(1:2)]          # species data to transform
x_mn <- min(dt[dt > 0])
x_mx <- max(dt)
rng <- 6           # abundance range
w <- log(rng) / (log(x_mx) - log(x_mn))
a <- x_mn^(-w)
# otu_6 <- a * dt[, -c(1:3)]^w
otu_6 <- a * dt^w
range(otu_6)
```

## Ordination methods
### DCA PA
```{r}
dca_pa <- decorana(veg = otu_pa)

print(dca_pa, head=T)
```


### GNMDS PA

#### Distances - don't run if loading saved object

```{r}
## Bray-Curtis
dist_pa <- vegdist(x = otu_pa, method = "bray")

## Geodist
ep <- 0.8     # epsilon
geodist_pa <- isomapdist(dist = dist_pa, epsilon = ep)
```


##### Save the result - don't run if loading saved object
```{r}
saveRDS(geodist_pa,
        file = (file.path(dataPath,"inputs/geodist_pa.rds")))
```

#### Ordination

#### monoMDS - don't run if loading saved object
```{r}

Sys.time()
d <- 2
mds_pa <- list()

for (i in 1:100) {
  mds_pa[[i]]<-monoMDS(geodist_pa,
                    matrix(c(runif(dim(otu_pa)[1]*d)),
                           nrow = dim(otu_pa)[1]),
                    k = d,
                    model = "global",
                    maxit = 2000,
                    smin = 1e-7,
                    sfgrmin = 1e-7)
}

```

##### Save the result - don't run if loading saved object
```{r}
saveRDS(mds_pa,
        file = file.path(dataPath,"inputs/mds_pa.rds"))
```


##### Best nmds solution - PA
Make function?
```{r, eval=TRUE}
# Loading geodist object
# geodist_nfi <- readRDS(file = "../SplitRev2_geodist_pa_full.rds")
```


```{r}
# Loading mds results
# mds <- readRDS(file = "../SplitRev2_mds_pa_full.rds")

## Extracting the stress of each nmds iteration
mds_stress_pa<-unlist(lapply(mds_pa, function(v){v[[22]]})) 

ordered_pa <-order(mds_stress_pa)

## Best, second best, and worst solution
mds_stress_pa[ordered_pa[1]]
mds_stress_pa[ordered_pa[2]]
mds_stress_pa[ordered_pa[10]]

## Scaling of axes to half change units and varimax rotation by postMDS
mds_best_pa<-postMDS(mds_pa[[ordered_pa[1]]],
                  geodist_pa, 
                  pc = TRUE, 
                  halfchange = TRUE, 
                  threshold = ep)     # Is this threshold related to the epsilon above?
mds_best_pa

mds_secbest_pa <- postMDS(mds_pa[[ordered_pa[2]]],
                          geodist_pa, 
                          pc = TRUE, 
                          halfchange = TRUE, 
                          threshold = ep)
mds_secbest_pa

## Procrustes comparisons
procr_pa <- procrustes(mds_best_pa,
                    mds_secbest_pa,
                    permutations=999)
protest(mds_best_pa,
        mds_secbest_pa,
        permutations=999)

plot(procr_pa)

png(file=file.path(dataPath,"outputs/procrustes_pa.png"), width=1000, height=700)
plot(procr_pa)
dev.off()
```

##### Correlation of axis: DCA vs NMDS - PA
```{r}
# Extracting ordination axis
ax <- 2
axis_pa <- cbind(mds_best_pa$points,
                 scores(dca_pa,
                        display = "sites",
                        origin = TRUE)[, 1:ax])

ggcorr(axis_pa, 
       method=c("everything","kendall"), 
       label = TRUE,
       label_size = 3, 
       label_color = "black",  
       nbreaks = 8,
       label_round = 3,
       low = "red",
       mid = "white",
       high = "green")


```

##### Save the figure
```{r}
ggsave(filename = file.path(dataPath,"outputs/correlationPCAvsNMDS_PA.png"),
       device = "png",
       dpi=300 )

# Switching direction of NMDS1
# mds_best$points[, 1] <- -mds_best$points[, 1]
```

### DCA R6
```{r}
dca_r6 <- decorana(veg = otu_6)

print(dca_r6, head=T)
```


### GNMDS R6
#### Distances - don't run if loading saved object
```{r}
## Bray-Curtis
dist_r6 <- vegdist(x = otu_6, method = "bray")

## Geodist
ep <- 0.80     # epsilon
geodist_r6 <- isomapdist(dist = dist_r6, epsilon = ep)
```

##### Save the result - don't run if loading saved object
```{r}
saveRDS(geodist_r6,
        file = (file.path(dataPath,"inputs/geodist_r6.rds")))
```


#### Ordination 

##### 200 reps - don't run if loading saved object
took 20mins
```{r}
# monoMDS
d <- 2
mds_r6 <- list()

Sys.time()
for (i in 1:200) {
  mds_r6[[i]]<-monoMDS(geodist_r6,
                    matrix(c(runif(dim(otu_6)[1]*d)),
                           nrow = dim(otu_6)[1]),
                    k = d,
                    model = "global",
                    maxit = 2000,
                    smin = 1e-7,
                    sfgrmin = 1e-7)
}
Sys.time()
```

##### Save the result - don't run if loading saved object
```{r}
saveRDS(mds_r6,
        file = file.path(dataPath,"inputs/mds_r6.rds")) 
```

##### Best nmds solution r6 200 rep 
```{r, eval=TRUE}
# Loading geodist object
# geodist_nfi <- readRDS(file = "../SplitRev2_geodist_nfi.rds")

# Loading mds results
# mds <- readRDS(file = "../SplitRev2_mds.rds")

## Extracting the stress of each nmds iteration
mds_stress_r6<-unlist(lapply(mds_r6, function(v){v[[22]]})) 

ordered_r6 <-order(mds_stress_r6)

## Best, second best, and worst solution
mds_stress_r6[ordered_r6[1]]
mds_stress_r6[ordered_r6[2]]
mds_stress_r6[ordered_r6[100]]

## Scaling of axes to half change units and varimax rotation by postMDS
mds_best_r6<-postMDS(mds_r6[[ordered_r6[1]]],
                     geodist_r6, 
                     pc = TRUE, 
                     halfchange = TRUE, 
                     threshold = ep)     # Is this threshold related to the epsilon above?
mds_best_r6

mds_secbest_r6<-postMDS(mds_r6[[ordered_r6[2]]],
                        geodist_r6, 
                        pc = TRUE, 
                        halfchange = TRUE, 
                        threshold = ep)
mds_secbest_r6

## Procrustes comparisons
procr_r6 <- procrustes(mds_best_r6,
                       mds_secbest_r6,
                       permutations=999)
protest(mds_best_r6,
        mds_secbest_r6,
        permutations=999)

plot(procr_r6)

png(file.path(dataPath,"outputs/procrustes_r6.png"), width=1000, height=700,) #added 1000
plot(procr_r6)
dev.off()
```

 #### 1000 reps - don't run if loading saved object 
 Currently commented out as it gained nothing but took extra time. Can be removed in due course.
<!-- took 1hr -->
<!-- ```{r} -->
<!-- # monoMDS -->
<!-- d <- 2 -->
<!-- mds_r6_1000 <- list()#obj added_1000 -->

<!-- Sys.time() -->
<!-- for (i in 1:1000) {#change from 200 (200 ra in 20min, 1000 ran in 1hr) -->
<!--   mds_r6_1000[[i]]<-monoMDS(geodist_r6,#obj added _1000 -->
<!--                     matrix(c(runif(dim(otu_6)[1]*d)), -->
<!--                            nrow = dim(otu_6)[1]), -->
<!--                     k = d, -->
<!--                     model = "global", -->
<!--                     maxit = 2000, -->
<!--                     smin = 1e-7, -->
<!--                     sfgrmin = 1e-7) -->
<!-- } -->
<!-- Sys.time() -->
<!-- ``` -->

<!-- ##### Save the result - don't run if loading saved object -->
<!-- ```{r} -->
<!-- saveRDS(mds_r6_1000, #obj added_1000 -->
<!--         file = file.path(dataPath,"inputs/mds_r6_1000rep.rds")) #change from 200 -->
<!-- ``` -->

<!-- ##### Best nmds solution r6 1000 rep -->
<!-- ```{r, eval=TRUE} -->
<!-- # Loading geodist object -->
<!-- # geodist_nfi <- readRDS(file = "../SplitRev2_geodist_nfi.rds") -->

<!-- # Loading mds results -->
<!-- # mds <- readRDS(file = "../SplitRev2_mds.rds") -->

<!-- ## Extracting the stress of each nmds iteration -->
<!-- mds_stress_r6_1000<-unlist(lapply(mds_r6_1000, function(v){v[[22]]})) #adeed 1000 -->

<!-- ordered_r6_1000 <-order(mds_stress_r6_1000) -->

<!-- ## Best, second best, and worst solution -->
<!-- mds_stress_r6_1000[ordered_r6_1000[1]] -->
<!-- mds_stress_r6_1000[ordered_r6_1000[2]] -->
<!-- mds_stress_r6_1000[ordered_r6_1000[100]] -->

<!-- ## Scaling of axes to half change units and varimax rotation by postMDS -->
<!-- mds_best_r6_1000<-postMDS(mds_r6_1000[[ordered_r6_1000[1]]], -->
<!--                      geodist_r6,  -->
<!--                      pc = TRUE,  -->
<!--                      halfchange = TRUE,  -->
<!--                      threshold = ep)     # Is this threshold related to the epsilon above? -->
<!-- mds_best_r6_1000 -->

<!-- mds_secbest_r6_1000<-postMDS(mds_r6_1000[[ordered_r6_1000[2]]], -->
<!--                         geodist_r6,  -->
<!--                         pc = TRUE,  -->
<!--                         halfchange = TRUE,  -->
<!--                         threshold = ep) -->
<!-- mds_secbest_r6_1000 -->

<!-- ## Procrustes comparisons -->
<!-- procr_r6_1000 <- procrustes(mds_best_r6_1000, -->
<!--                        mds_secbest_r6_1000, -->
<!--                        permutations=999) -->
<!-- protest(mds_best_r6_1000, -->
<!--         mds_secbest_r6_1000, -->
<!--         permutations=999) -->

<!-- plot(procr_r6_1000) -->

<!-- png(file.path(dataPath,"outputs/procrustes_r6_1000.png"), width=1000, height=700,) #added 1000 -->
<!-- plot(procr_r6_1000) -->
<!-- dev.off() -->
<!-- ``` -->



##### Correlation of axis: DCA vs NMDS 
retain the 200 rep version
```{r}
# Extracting ordination axis
ax <- 2
axis_r6 <- cbind(mds_best_r6$points,
                 scores(dca_r6,
                        display = "sites",
                        origin = TRUE)[, 1:ax])

ggcorr(axis_r6, 
       method=c("everything","kendall"), 
       label = TRUE,
       label_size = 3, 
       label_color = "black",  
       nbreaks = 8,
       label_round = 3,
       low = "red",
       mid = "white",
       high = "green")

```

##### Save the figure
```{r}
ggsave(filename = file.path(dataPath,"outputs/correlationPCAvsNMDS_r6.png"),
       device = "png",
       dpi=300 )

# Switching direction of NMDS1
# mds_best$points[, 1] <- -mds_best$points[, 1]
```

### Plotting DCA & GNMDS
```{r}
## Adding scores to data frame
otu_6$gnmds1 <- mds_best_r6$points[, 1]
otu_6$gnmds2 <- mds_best_r6$points[, 2]
otu_6$dca1 <- scores(dca_r6, display = "sites", origin = TRUE)[, 1]
otu_6$dca2 <- scores(dca_r6, display = "sites", origin = TRUE)[, 2]

p_gnmds_r6 <- ggplot(data = otu_6,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS",
          subtitle = "First run") +
  geom_point(colour = "red") +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")

p_dca_r6 <- ggplot(data = otu_6,
                   aes(x = dca1,
                       y = -dca2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("DCA",
          subtitle = "First run") +
  geom_point(colour = "red") +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")

p_gnmds_r6 + p_dca_r6


```

##### Save the figure
```{r}
ggsave(file.path(dataPath,"outputs/gnmds_dca_r6.png"),
       device = "png", 
       dpi=300)
```

## Species-environment relationships
### Selecting ordination
```{r}
ord <- mds_best_r6

## Axis scores if selected ord is GNMDS
axis <- ord$points %>% as.data.frame

## Axis scores if selected ord is DCA
# axis <- scores(ord,
#                display = "sites",
#                origin = TRUE)[, 1:ax])

```

### Create additional variables
decided to make MLD-bathy vars
```{r}
env<-env %>% 
  mutate ("MLDmean_bathy"=MLDmean_Robinson-(bathy*-1),
          "MLDmin_bathy"=MLDmin_Robinson-(bathy*-1),
          "MLDmax_bathy"=MLDmax_Robinson-(bathy*-1))

env$MLDmean_bathy<-cut(env$MLDmean_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))
env$MLDmin_bathy<-cut(env$MLDmin_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))
env$MLDmax_bathy<-cut(env$MLDmax_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))

env$swDensRob_avs<-swRho(salinity=env$Smean_Robinson,
                         temperature=env$Tmean_Robinson,
                         pressure=(env$bathy*-1),
                         eos="unesco")

```



### Correlation ordination axes and environmental variables
#### Removing non-env vars
```{r}
env_cont<-env%>% select(-c(landscape,sedclass,gmorph, MLDmean_bathy, MLDmax_bathy, MLDmin_bathy, #categorical
                           optional, #not a var
                           MLDmax_Robinson, MLDmean_Robinson, MLDmin_Robinson,  #replaced by new vars
                           MLDsd_Robinson #not meaningful
                           ))
env_cont<-env_cont%>% mutate_if(is.integer,as.numeric)
env_corr <- env_cont %>% select(-c(SampID))

# env_corr$coords.x1<-as.numeric(env_corr$coords.x1)
# env_corr$coords.x2<-as.numeric(env_corr$coords.x2)

env_corr[(!is.numeric(env_corr)),]
```

#### Correlations
```{r}
# Vector to hold correlations
cor_ax1 <- NULL
cor_ax2 <- NULL
pv_ax1 <- NULL
pv_ax2 <- NULL

# NMDS1
for( i in seq(length(env_corr))) {
  ct.i <- cor.test(axis$MDS1,
                   env_corr[, i],
                   method = "kendall")
  cor_ax1[i] <- ct.i$estimate
  pv_ax1[i] <- ct.i$p.value
}

# NMDS2
for( i in seq(length(env_corr))) {
  ct.i <- cor.test(axis$MDS2,
                   env_corr[, i],
                   method = "kendall")
  cor_ax2[i] <- ct.i$estimate
  pv_ax2[i] <- ct.i$p.value
}

cor_tab <- data.frame(env = names(env_corr),
                      ord_ax1 = cor_ax1,
                      pval_ax1 = pv_ax1,
                      ord_ax2 = cor_ax2,
                      pval_ax2 = pv_ax2)

cor_tab

write.csv(x = cor_tab,
          file = file.path(dataPath,"inputs/cor-table_r6_200rep_MLD-bathy.csv"),
          row.names = FALSE)
```
### Dot chart to check for gaps in correlation

```{r}
cor_a1_sort<-cor_tab%>%
  mutate(abs_ord_ax1=abs(ord_ax1),
         abs_ord_ax2=abs(ord_ax2)) %>%
  arrange(desc(abs_ord_ax1))

dotchart(cor_a1_sort$abs_ord_ax1, main="Absolute (+/-) correlations between envVars and gnmds axis 1")

cor_cut<-0.35 #decide

cor_sel<-subset(cor_a1_sort,abs_ord_ax1>cor_cut)
cor_sel
```


### Sel env var (top corr)
```{r}
env_os <- env[, cor_sel$env]
env_os
str(env_os)

```


### Ordisurfs top corr
```{r}
ordsrfs <- list(length = ncol(env_os))

for (i in seq(ncol(env_os))) {
  os.i <- gg_ordisurf(ord = ord,
                      env.var = env_os[, i],
                      pt.size = 1,
                      # binwidth = 0.05,
                      var.label = names(env_os)[i],
                      gen.text.size = 10,
                      title.text.size = 15,
                      leg.text.size = 10)
  
  ordsrfs[[i]] <- os.i$plot
}

ordsrfs_plt <- ggarrange(plotlist = ordsrfs,
                         nrow = 5,
                         ncol = 3)

ordsrfs_plt
```
##### Save some outputs
```{r}
ggexport(ordsrfs_plt,
          filename = file.path(dataPath,"outputs/ordisurfs_top_corr.png"),
          width = 2000,
          height = 3000)

```


### Sel env var (manual)
```{r}
env_os_m <- env[,c("Tmean_Robinson", #top corr
                  "salt_max", #top corr
                  "Smax_Robinson", #comparison to top corr
                  "swDensRob_avs", #top corr
                  "BO22_icecoverltmax_ss",#top corr ax2
                  "BO22_icecovermean_ss",#top corr
                  "BO22_dissoxmean_bdmean",#top corr
                  #"BO22_carbonphytoltmin_bdmean",#top corr - no clear gradient in ordisurf
                  "BO22_ppltmin_ss", #top corr
                  "X.y", #comparison to Y
                  "Y", #top corr
                  "spd_std", #top corr ax2 (blended model)
                  "CSpdsd_Robinson", #comparison to top corr ax2 (blended model)
                  "mud", #highest sed var ax1 + corr
                  "gravel",#highest sed var ax1 - corr
                  "BO22_silicateltmax_bdmean", #just under top corr ax1
                  "bathy" #intuitive for comparisons
                )]
env_os_m
str(env_os_m)

```


### Ordisurfs top corr
```{r}
ordsrfs_m <- list(length = ncol(env_os_m))

for (i in seq(ncol(env_os_m))) {
  os.i_m <- gg_ordisurf(ord = ord,
                      env.var = env_os_m[, i],
                      pt.size = 1,
                      # binwidth = 0.05,
                      var.label = names(env_os_m)[i],
                      gen.text.size = 10,
                      title.text.size = 15,
                      leg.text.size = 10)
  
  ordsrfs_m[[i]] <- os.i_m$plot
}

ordsrfs_plt_m <- ggarrange(plotlist = ordsrfs_m,
                         nrow = 4,
                         ncol = 4)

ordsrfs_plt_m
```

##### Save some outputs
```{r}
ggexport(ordsrfs_plt_m,
          filename = file.path(dataPath,"outputs/ordisurfs_man_sel.png"),
          width = 2000,
          height = 2000)

```




### Envfit
```{r}
## Select if any var should be excluded from envfit (makes less busy to read)
env_os_m_envfit<-env_os_m [,c("Tmean_Robinson", #top corr
                  "salt_max", #top corr
                  "Smax_Robinson", #comparison to top corr
                  "swDensRob_avs", #top corr
                  "BO22_icecoverltmax_ss",#top corr ax2
               #   "BO22_icecovermean_ss",#top corr
                  "BO22_dissoxmean_bdmean",#top corr
                  #"BO22_carbonphytoltmin_bdmean",#top corr - no clear gradient in ordisurf
                  "BO22_ppltmin_ss", #top corr
                  "X.y", #comparison to Y
                  "Y", #top corr
                  "spd_std", #top corr ax2 (blended model)
                  "CSpdsd_Robinson", #comparison to top corr ax2 (blended model)
                  "mud", #highest sed var ax1 + corr
                  "gravel",#highest sed var ax1 - corr
                  "BO22_silicateltmax_bdmean", #just under top corr ax1
                  "bathy" #intuitive for comparisons
                
  
)]

colnames(env_os_m_envfit)<-c("TmeanR", #top corr
                  "salt_max", #top corr
                  "SmaxR", #comparison to top corr
                  "swDensR", #top corr
                  "icecovmax",#top corr ax2
               #   "icecovav",#top corr
                  "dissoxav",#top corr
                  #"BO22_carbonphytoltmin_bdmean",#top corr - no clear gradient in ordisurf
                  "ppltmin", #top corr
                  "X", #comparison to Y
                  "Y", #top corr
                  "spd_std", #top corr ax2 (blended model)
                  "CSpdsdR", #comparison to top corr ax2 (blended model)
                  "mud", #highest sed var ax1 + corr
                  "gravel",#highest sed var ax1 - corr
                  "SiLtmax", #just under top corr ax1
                  "bathy" #intuitive for comparisons
                 )

## Envfot plot
gg_envfit(ord = ord,
          env = env_os_m_envfit,
          pt.size = 1)

## Envfit analysis

ef <- envfit(ord = ord,
             env = env_os_m_envfit,
             # na.rm = TRUE
             )
efDF <- as.data.frame(scores(ef,
                             display = "vectors"))
```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/EnvFit_man_sel_cln.png"),
       device = "png",
       dpi=300 )
```
### Finding Regional (biogeographic) Splits

#### add axis vals, sampNo to env_os_m

```{r}
env_sub_meta<-env_os_m

env_sub_meta$gnmds1 <- otu_6$gnmds1
env_sub_meta$gnmds2 <- otu_6$gnmds2
env_sub_meta$SampID <- otu_6$SampID
env_sub_meta$X <- env$X.y
```

#### Plot var vs var with gnmds ax1 colours
##### Bathy v Temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by gnmds r6 1000 rep axis 1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/bathy_v_temp_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Salinity v Temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Temp (mean Robinson) vs Salinity (max Robinson) - coloured by gnmds r6 1000 rep axis 1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/tempRmean_v_salinityRmax_nmds1.png"),
       device = "png",
       dpi=300 )
```


##### X v Y
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Geography (X v Y) - coloured by mean temp")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/X_v_Y_tmeanR.png"),
       device = "png",
       dpi=300 )
```
##### Temp v disoxltmin
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = BO22_dissoxltmin_bdmean)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Temp (mean Robinson) v disox lt min - coloured by gnmds axis 1")

tb_ax1

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/TmeanR_v_dissoxltmin_gnmds1.png"),
       device = "png",
       dpi=300 )
Sys.time()
```

<!-- ### Data split plot -->
<!-- #### Attaching ordination scores -->
<!-- ```{r} -->
<!-- env$gnmds1 <- axis$MDS1 -->
<!-- env$gnmds2 <- axis$MDS2 -->
<!-- ``` -->

<!-- #### Plots -->
<!-- ```{r} -->
<!-- TB_splitRaw <- ggplot(data = env, -->
<!--                       aes(x = temp_mean.tif, -->
<!--                           y = bathy.tif)) + -->
<!--   theme_classic() + -->
<!--   geom_point(aes(colour = SplitRaw), -->
<!--              size = 2) + -->
<!--   scale_colour_brewer(palette = "Paired") + -->
<!--   # geom_vline(xintercept = 0) + -->
<!--   # geom_segment(aes(x = 2, y = 0, -->
<!--   #                  xend = 2, yend = -675)) + -->
<!--   # geom_segment(aes(x = 0, y = -700, -->
<!--   #                  xend = 5, yend = -575)) + -->
<!--   # geom_segment(aes(x = 0, y = -700, -->
<!--   #                  xend = -1.5, yend = -700)) + -->
<!--   # geom_segment(aes(x = 0, y = -425, -->
<!--   #                  xend = 2, yend = -325)) + -->
<!--   # geom_segment(aes(x = 2, y = -500, -->
<!--   #                  xend = 5, yend = -400)) + -->
<!--   # geom_segment(aes(x = 5, y = -400, -->
<!--   #                  xend = 8, yend = 0)) + -->
<!--   # geom_segment(aes(x = 0, y = -425, -->
<!--   #                  xend = -1.5, yend = -425)) + -->
<!--   # annotate(geom = "text", x = -1, y = -1500, label = "1", size = 7) + -->
<!--   # annotate(geom = "text", x = -1, y = -75, label = "2", size = 7) + -->
<!--   # annotate(geom = "text", x = 2.5, y = -1000, label = "3", size = 7) + -->
<!--   # annotate(geom = "text", x = 1, y = -550, label = "4", size = 7) + -->
<!--   # annotate(geom = "text", x = 0.5, y = -250, label = "5", size = 7) + -->
<!--   # annotate(geom = "text", x = 6, y = -500, label = "6", size = 7) + -->
<!--   # annotate(geom = "text", x = 4.5, y = -250, label = "7", size = 7) + -->
<!--   # annotate(geom = "text", x = 1.1, y = 50, label = "8", size = 7) + -->
<!--   ggtitle("Temp (mean) vs Bathy - coloured by SplitRaw") -->

<!-- TB_split <- ggplot(data = env, -->
<!--                       aes(x = temp_mean.tif, -->
<!--                           y = bathy.tif)) + -->
<!--   theme_classic() + -->
<!--   geom_point(aes(colour = Split), -->
<!--              size = 2) + -->
<!--   scale_colour_brewer(palette = "Paired") + -->
<!--   # geom_vline(xintercept = 0) + -->
<!--   # geom_segment(aes(x = 2, y = 0, -->
<!--   #                  xend = 2, yend = -675)) + -->
<!--   # geom_segment(aes(x = 0, y = -700, -->
<!--   #                  xend = 5, yend = -575)) + -->
<!--   # geom_segment(aes(x = 0, y = -700, -->
<!--   #                  xend = -1.5, yend = -700)) + -->
<!--   # geom_segment(aes(x = 0, y = -425, -->
<!--   #                  xend = 2, yend = -325)) + -->
<!--   # geom_segment(aes(x = 2, y = -500, -->
<!--   #                  xend = 5, yend = -400)) + -->
<!--   # geom_segment(aes(x = 5, y = -400, -->
<!--   #                  xend = 8, yend = 0)) + -->
<!--   # geom_segment(aes(x = 0, y = -425, -->
<!--   #                  xend = -1.5, yend = -425)) + -->
<!--   # annotate(geom = "text", x = -1, y = -1500, label = "1", size = 7) + -->
<!--   # annotate(geom = "text", x = -1, y = -75, label = "2", size = 7) + -->
<!--   # annotate(geom = "text", x = 2.5, y = -1000, label = "3", size = 7) + -->
<!--   # annotate(geom = "text", x = 1, y = -550, label = "4", size = 7) + -->
<!--   # annotate(geom = "text", x = 0.5, y = -250, label = "5", size = 7) + -->
<!--   # annotate(geom = "text", x = 6, y = -500, label = "6", size = 7) + -->
<!--   # annotate(geom = "text", x = 4.5, y = -250, label = "7", size = 7) + -->
<!--   # annotate(geom = "text", x = 1.1, y = 50, label = "8", size = 7) + -->
<!--   ggtitle("Temp (mean) vs Bathy - coloured by Split") -->

<!-- EN_split <- ggplot(data = env, -->
<!--               aes(x = coords.x1, -->
<!--                   y = coords.x2)) + -->
<!--   theme_classic() + -->
<!--   geom_point(aes(colour = Split), -->
<!--              size = 2) + -->
<!--   scale_colour_brewer(palette = "Paired") + -->
<!--   ggtitle("Easting vs Northing - coloured by Split") -->

<!-- EN_split + TB_split -->



<!-- ``` -->

<!-- ### Coordinates coloured by GNMDS1 & 2 -->
<!-- ```{r} -->
<!-- EN1 <- ggplot(data = env, -->
<!--               aes(x = coords.x1, -->
<!--                   y = coords.x2)) + -->
<!--   theme_classic() + -->
<!--   geom_point(aes(colour = gnmds1), -->
<!--              size = 2) + -->
<!--   scale_colour_gradient2(low = "red", -->
<!--                          mid = "yellow", -->
<!--                          high = "green") + -->
<!--   ggtitle("Easting vs Northing - coloured by GNMDS1") -->

<!-- EN2 <- ggplot(data = env, -->
<!--               aes(x = coords.x1, -->
<!--                   y = coords.x2)) + -->
<!--   theme_classic() + -->
<!--   geom_point(aes(colour = gnmds2), -->
<!--              size = 2) + -->
<!--   scale_colour_gradient2(low = "red", -->
<!--                          mid = "yellow", -->
<!--                          high = "green") + -->
<!--   ggtitle("Easting vs Northing - coloured by GNMDS2") -->

<!-- EN_temp <- ggplot(data = env, -->
<!--               aes(x = coords.x1, -->
<!--                   y = coords.x2)) + -->
<!--   theme_classic() + -->
<!--   geom_point(aes(colour = temp_mean.tif), -->
<!--              size = 2) + -->
<!--   scale_colour_gradient2(low = "Blue", -->
<!--                          mid = "yellow", -->
<!--                          high = "red", -->
<!--                          midpoint = 4) + -->
<!--   ggtitle("Easting vs Northing - coloured by temp") -->

<!-- EN1 + EN2 -->
<!-- ``` -->


<!-- ## Convert to r script -->
<!-- ```{r} -->
<!-- library(knitr) -->
<!-- knitr::purl(input = "first-run.Rmd", -->
<!--             output = "first-run.r") -->
<!-- ``` -->

