---
title: "Splitting Data for NiN v 3 analyses - R Notebook"
output: html_notebook
---

#INPUTS
Assumes you hve already run at least the scripts:
- workflow_deepseaNiN_Start
- Workflow_deepseaNiN_Start_splitBiogeog

(Else further subdivisions of the same.) 

Requires the main analysis environment to be loaded.

# Selecting geographic subsets
This is to explore thresholds/relationships locally. E.g. To look for threshold values/ local relationships before deciding on splits. You need to already have a theory that you want to test in order to select the geographic regions to look at.

In this instance, for example, we are interested in at least making one split (into 2 groups) by temperature. We expect this to separate the shelf data from the combined deep/arctic data. We therefore should have a look at at least 3 areas: 
- the slope off lofoten (Norland VI/Norland VII) to see the temperature gradient with depth
- the finmark coast from Tromsøflaket to the southern end of Mareano Øst to follow the temperature gradient east into the Barents sea
- the MAREANO Øst region where we can follow a temperature gradient from the coast North into arctic water

### Choose area:
Choosing the exact areas can be done using the interactive "ggplotly" XY map of temperature to zoom in on the area concerned and note the bounding box values to use.

That is reproduced here:
##### X v Y - temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-2, 10),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp")

ggplotly(tb_ax1)
```

This gives us the following bounding box selections

NorlandSlope: X:205261 - 364461, Y: 743294 - 7621734
FinnmarkCoast: X: 599461 - 1138061, Y: 7794534 - 8013534
MareanoØst: X: 1009861 - 1162261, Y: 7834134 - 8505934

Additional selections made after investigating these three areas were made to check the temperature thrshold on the slope in the north and south:

SvalbardSlope: X: 361061 - 561862, Y: 7915734 - 8782734
MøreSlope: X: -47738 - 93861, Y: 6976734 - 7234534

After these explorations it was decided to generate a subset by density. A threshold of 1031.8 was used to split the dataset.

## Make data subsets to explore:

```{r}
NorlandSlopeJoinedDat<-subset(joinedDat, X.y>205261 & X.y<364461 & Y>743294 & Y<7621734)
plot(NorlandSlopeJoinedDat$X.y, NorlandSlopeJoinedDat$Y)
NSotu_sort<-NorlandSlopeJoinedDat %>% select(c(2:351))
NSenv_sort<-NorlandSlopeJoinedDat %>% select(c(2,352:491))

FinnmarkCoastJoinedDat<-subset(joinedDat, X.y>599461 & X.y<1138061 & Y>7794534 & Y<8013534)
plot(FinnmarkCoastJoinedDat$X.y, FinnmarkCoastJoinedDat$Y)
FCotu_sort<-FinnmarkCoastJoinedDat %>% select(c(2:351))
FCenv_sort<-FinnmarkCoastJoinedDat %>% select(c(2,352:491))

MareanoØstJoinedDat<-subset(joinedDat, X.y>1009861 & X.y<1162261 & Y>7834134 & Y<8505934)
plot(MareanoØstJoinedDat$X.y, MareanoØstJoinedDat$Y)
MØotu_sort<-MareanoØstJoinedDat %>% select(c(2:351))
MØenv_sort<-MareanoØstJoinedDat %>% select(c(2,352:491))

### Second selections

SvalbardSlopeJoinedDat<-subset(joinedDat, X.y>361061 & X.y<561862 & Y>7915734 & Y<8782734)
plot(SvalbardSlopeJoinedDat$X.y, SvalbardSlopeJoinedDat$Y)
SSotu_sort<-SvalbardSlopeJoinedDat %>% select(c(2:351))
SSenv_sort<-SvalbardSlopeJoinedDat %>% select(c(2,352:491))

MøreSlopeJoinedDat<-subset(joinedDat, X.y>-47738 & X.y<93861 & Y>6976734 & Y<7234534)
plot(MøreSlopeJoinedDat$X.y, MøreSlopeJoinedDat$Y)
MSotu_sort<-MøreSlopeJoinedDat %>% select(c(2:351))
MSenv_sort<-MøreSlopeJoinedDat %>% select(c(2,352:491))

#density split 1031.8 threshold:

joinedDat$swDensRob_avs<-swRho(salinity=joinedDat$Smean_Robinson,
                         temperature=joinedDat$Tmean_Robinson,
                         pressure=(joinedDat$bathy*-1),
                         eos="unesco")

loDensShallowJoinedDat<-subset(joinedDat,swDensRob_avs<1031.8)
plot(loDensShallowJoinedDat$X.y, loDensShallowJoinedDat$Y)
LDensShlw_otu_sort<-loDensShallowJoinedDat %>% select(c(2:351))
LDensShlw_env_sort<-loDensShallowJoinedDat %>% select(c(2,352:491))                                


hiDensDeepJoinedDat<-subset(joinedDat,swDensRob_avs>1031.8)
plot(hiDensDeepJoinedDat$X.y, hiDensDeepJoinedDat$Y)
HDensDeep_otu_sort<-hiDensDeepJoinedDat %>% select(c(2:351))
HDensDeep_env_sort<-hiDensDeepJoinedDat %>% select(c(2,352:491))  


```
###Save datasets
```{r}
write.csv(NSotu_sort, file=file.path(dataPath,(paste0("inputs/NSotu_sort_", Sys.Date(),".csv"))))
write.csv(NSenv_sort, file=file.path(dataPath,(paste0("inputs/NSenv_sort_", Sys.Date(),".csv"))))

write.csv(FCotu_sort, file=file.path(dataPath,(paste0("inputs/FCotu_sort_", Sys.Date(),".csv"))))
write.csv(FCenv_sort, file=file.path(dataPath,(paste0("inputs/FCenv_sort_", Sys.Date(),".csv"))))

write.csv(MØotu_sort, file=file.path(dataPath,(paste0("inputs/MOotu_sort_", Sys.Date(),".csv"))))
write.csv(MØenv_sort, file=file.path(dataPath,(paste0("inputs/MOenv_sort_", Sys.Date(),".csv"))))

### second set

write.csv(SSotu_sort, file=file.path(dataPath,(paste0("inputs/SSotu_sort_", Sys.Date(),".csv"))))
write.csv(SSenv_sort, file=file.path(dataPath,(paste0("inputs/SSenv_sort_", Sys.Date(),".csv"))))

write.csv(MSotu_sort, file=file.path(dataPath,(paste0("inputs/MSotu_sort_", Sys.Date(),".csv"))))
write.csv(MSenv_sort, file=file.path(dataPath,(paste0("inputs/MSenv_sort_", Sys.Date(),".csv"))))

### density splits

write.csv(HDensDeep_otu_sort, file=file.path(dataPath,(paste0("inputs/HDensDeep_otu_sort_", Sys.Date(),".csv"))))
write.csv(HDensDeep_env_sort, file=file.path(dataPath,(paste0("inputs/HDensDeep_env_sort_", Sys.Date(),".csv"))))

write.csv(LDensShlw_otu_sort, file=file.path(dataPath,(paste0("inputs/LDensShlw_otu_sort_", Sys.Date(),".csv"))))
write.csv(LDensShlw_env_sort, file=file.path(dataPath,(paste0("inputs/LDensShlw_env_sort_", Sys.Date(),".csv"))))

```
