---
title: "MAREANO - NiN:splitting biogeographic regions"
output: html_notebook
---

Asssumes you have aready run:
- config
- workflow_deepseaNiN_Start

You now have the objects needed to explore the dataset and find the best way to apply biogeographic splits. This is a pre-step before running the workflow again on the split dataset to explore possible LKMs.

# Finding Regional (biogeographic) Splits

This requires exploring the dataset to theorise the best ways to split the data with:
- environmental biplots/colourings


then applying the splits. 


## Exploring biplots
#### add axis vals, sampNo to env_os_m

```{r}
env_sub_meta<-env_os_m

env_sub_meta$gnmds1 <- otu_6$gnmds1
env_sub_meta$gnmds2 <- otu_6$gnmds2
env_sub_meta$SampID <- otu_6$SampID
env_sub_meta$X <- env$X.y
```
### Plots with continuous colour scales (variables)



#### Plot var vs var with gnmds ax1 colours
##### Bathy v Temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
  scale_colour_gradientn(limits = c(-3, 3),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by gnmds r6 200 rep axis 1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/bathy_v_temp_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Bathy v Salinity coloured by temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-2, 10),
                        colors=c('red','yellow','green'))+
  ggtitle("Salinity (max Robinson) vs Bathy - coloured by av Temp (R)")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/bathy_v_salMax_temp.png"),
       device = "png",
       dpi=300 )
```

##### Salinity v Temp gnmds1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
scale_colour_gradientn(limits = c(-3, 3),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Salinity (max R) - coloured by gnmds r6 200 rep axis 1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/tempRmean_v_salinityRmax_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Salinity v Temp dissoxmean
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = BO22_dissoxmean_bdmean),
             size = 1) +
scale_colour_gradientn(limits = c(285, 375),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Salinity (max R) - coloured by dissolved oxygen")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/tempRmean_v_salinityRmax_disooxmean.png"),
       device = "png",
       dpi=300 )
```

##### X v Y - temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-2, 10),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/X_v_Y_tmeanR.png"),
       device = "png",
       dpi=300 )
```

##### X v Y - density
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = swDensRob_avs),
             size = 1) +
  scale_colour_gradientn(limits = c(1027, 1041),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by water density")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/X_v_Y_swDenRobavs.png"),
       device = "png",
       dpi=800 )
```

##### Temp v disoxltmin
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = BO22_dissoxmean_bdmean)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
  scale_colour_gradientn(limits = c(-3, 3),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) v disox mean - coloured by gnmds axis 1")

tb_ax1

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/TmeanR_v_dissoxltmin_gnmds1.png"),
       device = "png",
       dpi=300 )

```
