---
title: "MAREANO - NiN:splitting biogeographic regions_BarentsDeep"
output: html_notebook
---

Asssumes you have aready run:
- config
- workflow_deepseaNiN_Start

You now have the objects needed to explore the dataset and find the best way to apply biogeographic splits. This is a pre-step before running the workflow again on the split dataset to explore possible LKMs.

# Finding Regional (biogeographic) Splits

This requires exploring the dataset to theorise the best ways to split the data with:
- environmental biplots/colourings


then applying the splits. 


### Libraries
```{r}
library(plotly)
library(readxl)
```




## Exploring biplots
#### make full dataset to explore 
should include all env Vars and otus to explore, and make species richness variable

```{r}
#add species data and sp richness variable
env_sub_meta1<-left_join(env,otu_6, by="SampID")
env_sub_meta1$spRich<-rowSums(otu_6[,-c(1:which(colnames(otu_6)=="Zoanthidae"))]!=0)

#rename X
env_sub_meta1$X <- env$X.y
env_sub_meta1 <-env_sub_meta1 %>% select (-c(X.y))

#add samplID
#env_sub_meta1$SampID<-envSel$SampID

```


### Add provisional biotope data to env file
Note that biotopes were last assigned in march 2022 and therefore there are some addional samples that have not yet got a biotope assigned. These should just be NAs
```{r}
biotopeInfo<-read_xlsx(file.path(dataPath, "inputs/MAREANO_provisional_biotope_classification_0322.xlsx"), sheet=1) %>%
  select(-c(x_coordinate_UTM33N, y_coordinate_UTM33N))

env_sub_meta<-left_join(env_sub_meta1,biotopeInfo)
```


### a colour palette for discrete plots (see later in script)

```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

### Find min max of vars for scales
```{r}
summary(env_sub_meta)
```



### Plots with CONTINUOUS colour scales (variables)


##### Bathy v Temp w gnmds ax 1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
  scale_colour_gradientn(limits = c(-1.8, 1.9),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by gnmds r6 ax1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeepbathy_v_temp_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Bathy v Temp w salinity
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Smax_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(34.43, 35.43),
                        colors=c('green','yellow','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by salinity")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeepbathy_v_temp_salinityMax.png"),
       device = "png",
       dpi=300 )
```



##### Bathy v Temp w longitude
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = X.y),
             size = 1) +
  scale_colour_gradientn(limits = c(-59139, 1168661),
                        colors=c('red','yellow','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by longitude")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeepbathy_v_temp_longitude.png"),
       device = "png",
       dpi=300 )
```
##### Bathy v Temp w latitude
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Y),
             size = 1) +
  scale_colour_gradientn(limits = c(7015934, 8949334),
                        colors=c('red','yellow','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by latitude")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeepbathy_v_temp_latitude.png"),
       device = "png",
       dpi=300 )
```
##### Bathy v Temp w iceCovLTMax
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = BO22_icecoverltmax_ss),
             size = 1) +
  scale_colour_gradientn(limits = c(0,0.93),
                        colors=c('grey','turquoise','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by ice cover LT max")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeepbathy_v_temp_iceCovLTmax.png"),
       device = "png",
       dpi=300 )
```


##### Bathy v Salinity coloured by temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-1.31, 6.68),
                        colors=c('red','yellow','green'))+
  ggtitle("Salinity (max Robinson) vs Bathy - coloured by av Temp (R)")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeepbathy_v_salMax_temp.png"),
       device = "png",
       dpi=300 )
```

##### Salinity v Temp gnmds1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
scale_colour_gradientn(limits = c(-2, 2.5),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Salinity (max R) - coloured by gnmds r6 ax 1 - grey 2.5-5")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeeptempRmean_v_salinityRmax_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Salinity v Temp dissoxmean
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = BO22_dissoxmean_bdmean),
             size = 1) +
scale_colour_gradientn(limits = c(282.5, 372.2),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Salinity (max R) - coloured by dissolved oxygen")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeeptempRmean_v_salinityRmax_disooxmean.png"),
       device = "png",
       dpi=300 )
```
##### TS - icecoverLTMax
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = BO22_icecoverltmax_ss),
             size = 1) +
  scale_colour_gradientn(limits = c(0,0.93),
                        colors=c('blue','green','red'))+
  ggtitle("Temp (AvR) v Salinity (maxR) - coloured by ice cover LT max")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeeptempRmean_v_salinityRmax_iceCoveLTmax.png"),
       device = "png",
       dpi=800 )
```


##### X v Y - temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-1.1, 8.5),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp")

ggplotly(tb_ax1)

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_X_v_Y_tmeanR.png"),
       device = "png",
       dpi=300 )
```

##### X v Y - density
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = swDensRob_avs),
             size = 1) +
  scale_colour_gradientn(limits = c(1027, 1031),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by water density")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_X_v_Y_swDenRobavs.png"),
       device = "png",
       dpi=800 )
```
##### X v Y - dissox
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = BO22_dissoxmean_bdmean),
             size = 1) +
  scale_colour_gradientn(limits = c(282,373),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by dissovled oxygen")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_X_v_Y_dissox.png"),
       device = "png",
       dpi=800 )
```
##### X v Y - icecoverLTMax
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = BO22_icecoverltmax_ss),
             size = 1) +
  scale_colour_gradientn(limits = c(0.00001,0.93),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by ice cover LT max - grey <0.00001")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_X_v_Y_iceCoveLTmax.png"),
       device = "png",
       dpi=800 )
```


##### Temp v disoxltmin gnmds
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = BO22_dissoxmean_bdmean)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
  scale_colour_gradientn(limits = c(-2.3, 2),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) v disox mean - coloured by gnmds ax 1")

tb_ax1

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_TmeanR_v_dissoxltmin_gnmds1.png"),
       device = "png",
       dpi=300 )

```
##### Temp v disoxltmin bathy
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = BO22_dissoxmean_bdmean)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 1) +
  scale_colour_gradientn(limits = c(-702, -38),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) v disox mean - coloured by bathymetry")

tb_ax1

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeepTmeanR_v_dissoxltmin_bathy.png"),
       device = "png",
       dpi=300 )

```



### Plots with DISCRETE colour scales (variables)

#### X v Y with mld-bathy categories
```{r}
dis_split <- ggplot(data = env,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = MLDmean_bathy),
             size = 1) +
  scale_colour_manual(values=cbPalette)+
 # scale_colour_brewer(palette = "Set3") +
  ggtitle("Easting vs Northing - coloured by Mixed layer depth proximity")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_XvY_MLDmeanBathy.png"),
       device = "png",
       dpi=300 )

```

### categorise gnmds

```{r}
env_sub_meta$ax1cat<-cut(env_sub_meta$gnmds1, 
      breaks=c(-3.2,-3,-2,-1,0,1,2,3,3.46))

env_sub_meta$ax2cat<-cut(env_sub_meta$gnmds2, 
      breaks=c(-1.9,-1,0,1,2,3,4,4.9))
```


#### X v Y with gnmds ax 1 as HC categories
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = ax1cat),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by gnmds axis 1 HC units")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_XvY_ax1cat.png"),
       device = "png",
       dpi=300 )

```

### categorise temp 5.1

```{r}
env_sub_meta$temp5_1<-cut(env_sub_meta$Tmean_Robinson, 
      breaks=c(-1.1, 5.1, 8.5))
#
```


#### X v Y with temp 5.1
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = temp5_1),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Set1") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by temp thresholded at 5.1*C")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_XvY_temp5_1.png"),
       device = "png",
       dpi=300 )

```
### categorise dissox

```{r}
env_sub_meta$dissoxav305<-cut(env_sub_meta$BO22_dissoxmean_bdmean, 
     # breaks=c(256, 282, 360)) #ltmin
     breaks=c(282.5,305,372.2),
     labels=c("lowO2","hiO2")) #mean
#
```


#### X v Y with dissox
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = dissoxav305),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Set1") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by dissox av thresholded at 305")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_XvY_dissoxav305.png"),
       device = "png",
       dpi=300 )

```

### categorise iceCovLTMax

```{r}
env_sub_meta$iceMx_gt0<-cut(env_sub_meta$BO22_icecoverltmax_ss, 
      breaks=c(0, 0.00001, 0.92838))
#
```


#### X v Y with temp 5.1
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = iceMx_gt0),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Set1") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by ice LT Max thresholded at >0")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_XvY_iceLTmx_gt0.png"),
       device = "png",
       dpi=300 )

```




#### X v Y with gnmds ax 2 as HC categories
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = ax2cat),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by gnmds axis 2 HC units")

dis_split

```




##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_XvY_ax2cat.png"),
       device = "png",
       dpi=300 )

```

#### Gnmds w temp 5.1
```{r}

t_gmo <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by temp 5.1 threshold",
          subtitle = "First run") +
  geom_point(aes(colour = factor(temp5_5))) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

t_gmo
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_gnmds_temp5_1.png"),
       device = "png",
       dpi=300 )
```
#### Gnmds w dissox 305
```{r}

o_gmo <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by dissox 305 threshold",
          subtitle = "First run") +
  geom_point(aes(colour = factor(dissox305))) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

o_gmo
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_gnmds_dissox305.png"),
       device = "png",
       dpi=300 )
```

#### Gnmds w iceLT max >0
```{r}

i_gmo <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by ice max >0 threshold",
          subtitle = "First run") +
  geom_point(aes(colour = factor(iceMx_gt0))) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

i_gmo
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/BarentsDeep_gnmds_iceMxgt0.png"),
       device = "png",
       dpi=300 )
```

```{r}
comp<-i_gmo+o_gmo+t_gmo
```
##### Save the plot
```{r}


##### Save some outputs

ggexport(comp,
          filename = file.path(dataPath,"outputs/BarentsDeep_gnmds_compare.png"),
          width = 1500,
          height = 500)

```



# Save environment

EDIT AREA FIRST!

```{r}
save.image(file.path(dataPath,"inputs/Renv_deepseaNiN_BarentsDeep.RData")) # edit area first

```

NOW CLEAR ENVIRONMENT BEFORE RUNNING FOR NEW AREA

