---
title: "MAREANO - NiN:splitting biogeographic regions"
output: html_notebook
---

Asssumes you have aready run:
- config
- workflow_deepseaNiN_Start

You now have the objects needed to explore the dataset and find the best way to apply biogeographic splits. This is a pre-step before running the workflow again on the split dataset to explore possible LKMs.

# Finding Regional (biogeographic) Splits

This requires exploring the dataset to theorise the best ways to split the data with:
- environmental biplots/colourings


then applying the splits. 

### Libraries
```{r}
library(plotly)
```



## Exploring biplots
#### add axis vals, sampNo to env_os_m

```{r}
env_sub_meta<-env_os_m

env_sub_meta$gnmds1 <- otu_6$gnmds1
env_sub_meta$gnmds2 <- otu_6$gnmds2
env_sub_meta$SampID <- otu_6$SampID
env_sub_meta$X <- env$X.y
```
### a colour palette for discrete plots (see later in script)

```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


### Plots with CONTINUOUS colour scales (variables)


##### Bathy v Temp w gnmds ax 1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
  scale_colour_gradientn(limits = c(-3, 3),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by gnmds r6 200 rep axis 1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/bathy_v_temp_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Bathy v Temp w salinity
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Smax_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(34.4, 35.45),
                        colors=c('green','yellow','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by salinity")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/bathy_v_temp_salinityMax.png"),
       device = "png",
       dpi=300 )
```

##### Bathy v Temp w latitude
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Y),
             size = 1) +
  scale_colour_gradientn(limits = c(6944134, 8949734),
                        colors=c('red','yellow','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by latitude")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/bathy_v_temp_latitude.png"),
       device = "png",
       dpi=300 )
```
##### Bathy v Temp w longitude
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = X.y),
             size = 1) +
  scale_colour_gradientn(limits = c(-107938, 1162261),
                        colors=c('red','yellow','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by longitude")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/bathy_v_temp_longitude.png"),
       device = "png",
       dpi=300 )
```

##### Bathy v Salinity coloured by temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-2, 10),
                        colors=c('red','yellow','green'))+
  ggtitle("Salinity (max Robinson) vs Bathy - coloured by av Temp (R)")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/bathy_v_salMax_temp.png"),
       device = "png",
       dpi=300 )
```

##### Salinity v Temp gnmds1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
scale_colour_gradientn(limits = c(-3, 3),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Salinity (max R) - coloured by gnmds r6 200 rep axis 1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/tempRmean_v_salinityRmax_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Salinity v Temp dissoxmean
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = BO22_dissoxmean_bdmean),
             size = 1) +
scale_colour_gradientn(limits = c(285, 375),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Salinity (max R) - coloured by dissolved oxygen")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/tempRmean_v_salinityRmax_disooxmean.png"),
       device = "png",
       dpi=300 )
```

##### X v Y - temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-2, 10),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp")

ggplotly(tb_ax1)

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/X_v_Y_tmeanR.png"),
       device = "png",
       dpi=300 )
```

##### X v Y - density
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = swDensRob_avs),
             size = 1) +
  scale_colour_gradientn(limits = c(1027, 1041),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by water density")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/X_v_Y_swDenRobavs.png"),
       device = "png",
       dpi=800 )
```

##### Temp v disoxltmin gnmds
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = BO22_dissoxmean_bdmean)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
  scale_colour_gradientn(limits = c(-3, 3),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) v disox mean - coloured by gnmds axis 1")

tb_ax1

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/TmeanR_v_dissoxltmin_gnmds1.png"),
       device = "png",
       dpi=300 )

```
##### Temp v disoxltmin gnmds
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = BO22_dissoxmean_bdmean)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 1) +
  scale_colour_gradientn(limits = c(-2750, 0),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) v disox mean - coloured by bathymetry")

tb_ax1

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/TmeanR_v_dissoxltmin_bathy.png"),
       device = "png",
       dpi=300 )

```



### Plots with DISCRETE colour scales (variables)

#### X v Y with mld-bathy categories
```{r}
dis_split <- ggplot(data = env,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = MLDmean_bathy),
             size = 1) +
  scale_colour_manual(values=cbPalette)+
 # scale_colour_brewer(palette = "Set3") +
  ggtitle("Easting vs Northing - coloured by Mixed layer depth proximity")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/XvY_MLDmeanBathy.png"),
       device = "png",
       dpi=300 )

```

### categorise gnmds

```{r}
env_sub_meta$ax1cat<-cut(env_sub_meta$gnmds1, 
      breaks=c(-3.2,-3,-2,-1,0,1,2,3,3.46))

env_sub_meta$ax2cat<-cut(env_sub_meta$gnmds2, 
      breaks=c(-1.9,-1,0,1,2,3,4,4.9))
```


#### X v Y with gnmds ax 1 as HC categories
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = ax1cat),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by gnmds axis 1 HC units")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/XvY_ax1cat.png"),
       device = "png",
       dpi=300 )

```

#### X v Y with gnmds ax 2 as HC categories
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = ax2cat),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by gnmds axis 2 HC units")

dis_split

```




##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/XvY_ax2cat.png"),
       device = "png",
       dpi=300 )

```


# Selecting geographic subsets
This is to explore thresholds/relationships locally. E.g. To look for threshold values/ local relationships before deciding on splits. You need to already have a theory that you want to test in order to select the geographic regions to look at.

In this instance, for example, we are interested in at least making one split (into 2 groups) by temperature. We expect this to separate the shelf data from the combined deep/arctic data. We therefore should have a look at at least 3 areas: 
- the slope off lofoten (Norland VI/Norland VII) to see the temperature gradient with depth
- the finmark coast from Tromsøflaket to the southern end of Mareano Øst to follow the temperature gradient east into the Barents sea
- the MAREANO Øst region where we can follow a temperature gradient from the coast North into arctic water

### Choose area:
Choosing the exact areas can be done using the interactive "ggplotly" XY map of temperature to zoom in on the area concerned and note the bounding box values to use.

That is reproduced here:
##### X v Y - temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-2, 10),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp")

ggplotly(tb_ax1)
```

This gives us the following bounding box selections

NorlandSlope: X:205261 - 364461, Y: 743294 - 7621734
FinnmarkCoast: X: 599461 - 1138061, Y: 7794534 - 8013534
MareanoØst: X: 1009861 - 1162261, Y: 7834134 - 8505934

## Make data subsets to explore:

```{r}
NorlandSlopeJoinedDat<-subset(joinedDat, X.y>205261 & X.y<364461 & Y>743294 & Y<7621734)
plot(NorlandSlopeJoinedDat$X.y, NorlandSlopeJoinedDat$Y)
NSotu_sort<-NorlandSlopeJoinedDat %>% select(c(2:351))
NSenv_sort<-NorlandSlopeJoinedDat %>% select(c(2,352:491))

FinnmarkCoastJoinedDat<-subset(joinedDat, X.y>599461 & X.y<1138061 & Y>7794534 & Y<8013534)
plot(FinnmarkCoastJoinedDat$X.y, FinnmarkCoastJoinedDat$Y)
FCotu_sort<-FinnmarkCoastJoinedDat %>% select(c(2:351))
FCenv_sort<-FinnmarkCoastJoinedDat %>% select(c(2,352:491))

MareanoØstJoinedDat<-subset(joinedDat, X.y>1009861 & X.y<1162261 & Y>7834134 & Y<8505934)
plot(MareanoØstJoinedDat$X.y, MareanoØstJoinedDat$Y)
MØotu_sort<-MareanoØstJoinedDat %>% select(c(2:351))
MØenv_sort<-MareanoØstJoinedDat %>% select(c(2,352:491))
```








