---
title: "MAREANO - NiN:splitting biogeographic regions"
output: html_notebook
---

Asssumes you have aready run:
- config
- workflow_deepseaNiN_Start

You now have the objects needed to explore the dataset and find the best way to apply biogeographic splits. This is a pre-step before running the workflow again on the split dataset to explore possible LKMs.

# Finding Regional (biogeographic) Splits

This requires exploring the dataset to theorise the best ways to split the data with:
- environmental biplots/colourings


then applying the splits. 

### Libraries
```{r}
library(plotly)
```



## Exploring biplots
#### add axis vals, sampNo to env_os_m

```{r}
env_sub_meta<-env_os_m

env_sub_meta$gnmds1 <- otu_6$gnmds1
env_sub_meta$gnmds2 <- otu_6$gnmds2
env_sub_meta$SampID <- otu_6$SampID
env_sub_meta$X <- env$X.y
```
### a colour palette for discrete plots (see later in script)

```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```


### Plots with CONTINUOUS colour scales (variables)


##### Bathy v Temp w gnmds ax 1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
  scale_colour_gradientn(limits = c(-2.1, 2.2),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by gnmds r6 200 rep axis 1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NSbathy_v_temp_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Bathy v Temp w salinity
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Smax_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(34.6, 35.3),
                        colors=c('green','yellow','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by salinity")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NSbathy_v_temp_salinityMax.png"),
       device = "png",
       dpi=300 )
```



##### Bathy v Temp w longitude
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = X.y),
             size = 1) +
  scale_colour_gradientn(limits = c(599461, 1138061),
                        colors=c('red','yellow','blue'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by longitude")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NSbathy_v_temp_longitude.png"),
       device = "png",
       dpi=300 )
```

##### Bathy v Salinity coloured by temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = bathy)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(3.9, 7.1),
                        colors=c('red','yellow','green'))+
  ggtitle("Salinity (max Robinson) vs Bathy - coloured by av Temp (R)")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NSbathy_v_salMax_temp.png"),
       device = "png",
       dpi=300 )
```

##### Salinity v Temp gnmds1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
scale_colour_gradientn(limits = c(-2.1, 2.2),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Salinity (max R) - coloured by gnmds r6 200 rep axis 1")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NStempRmean_v_salinityRmax_nmds1.png"),
       device = "png",
       dpi=300 )
```
##### Salinity v Temp dissoxmean
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Smax_Robinson,
                          y = Tmean_Robinson)) +
  theme_classic() +
  geom_point(aes(colour = BO22_dissoxmean_bdmean),
             size = 1) +
scale_colour_gradientn(limits = c(294, 308),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Salinity (max R) - coloured by dissolved oxygen")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NStempRmean_v_salinityRmax_disooxmean.png"),
       device = "png",
       dpi=300 )
```

##### X v Y - temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(4, 7),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp")

ggplotly(tb_ax1)

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NS_X_v_Y_tmeanR.png"),
       device = "png",
       dpi=300 )
```

##### X v Y - density
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = swDensRob_avs),
             size = 1) +
  scale_colour_gradientn(limits = c(1027.3, 1029.7),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by water density")

tb_ax1

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NS_X_v_Y_swDenRobavs.png"),
       device = "png",
       dpi=800 )
```

##### Temp v disoxltmin gnmds
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = BO22_dissoxmean_bdmean)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 1) +
  scale_colour_gradientn(limits = c(-2.1, 2.2),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) v disox mean - coloured by gnmds axis 1")

tb_ax1

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NS_TmeanR_v_dissoxltmin_gnmds1.png"),
       device = "png",
       dpi=300 )

```
##### Temp v disoxltmin bathy
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson,
                          y = BO22_dissoxmean_bdmean)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 1) +
  scale_colour_gradientn(limits = c(-450, -35),
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) v disox mean - coloured by bathymetry")

tb_ax1

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NSTmeanR_v_dissoxltmin_bathy.png"),
       device = "png",
       dpi=300 )

```



### Plots with DISCRETE colour scales (variables)

#### X v Y with mld-bathy categories
```{r}
dis_split <- ggplot(data = env,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = MLDmean_bathy),
             size = 1) +
  scale_colour_manual(values=cbPalette)+
 # scale_colour_brewer(palette = "Set3") +
  ggtitle("Easting vs Northing - coloured by Mixed layer depth proximity")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NS_XvY_MLDmeanBathy.png"),
       device = "png",
       dpi=300 )

```

### categorise gnmds

```{r}
env_sub_meta$ax1cat<-cut(env_sub_meta$gnmds1, 
      breaks=c(-3.2,-3,-2,-1,0,1,2,3,3.46))

env_sub_meta$ax2cat<-cut(env_sub_meta$gnmds2, 
      breaks=c(-1.9,-1,0,1,2,3,4,4.9))
```


#### X v Y with gnmds ax 1 as HC categories
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = ax1cat),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by gnmds axis 1 HC units")

dis_split

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NS_XvY_ax1cat.png"),
       device = "png",
       dpi=300 )

```

#### X v Y with gnmds ax 2 as HC categories
```{r}
dis_split <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = ax2cat),
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by gnmds axis 2 HC units")

dis_split

```




##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/NS_XvY_ax2cat.png"),
       device = "png",
       dpi=300 )

```

# Save environment

EDIT AREA FIRST!

```{r}
save.image("I:/Scripts/deepseaNiN/Renv_deepseaNiN_NS.RData") # edit area first

```

NOW CLEAR ENVIRONMENT BEFORE RUNNING FOR NEW AREA

