---
title: "MAREANO - NiN:splitting biogeographic regions_Deep"
output: html_notebook
---
**NB THESE ARE ALL EXAMPLES OF PLOTS TO SPEED UP YOUR CUSTOM PLOT MAKING: NO NEED TO RUN ALL - JUST USE THIS SCRIPT AS A TEMPLATE**

Asssumes you have aready run:
- config
- workflow_deepseaNiN_Start

You now have the objects needed to explore the dataset and find the best way to apply biogeographic splits. This is a pre-step before running the workflow again on the split dataset to explore possible LKMs.

# Finding Regional (biogeographic) Splits

This requires exploring the dataset to theorise the best ways to split the data with:
- environmental biplots/colourings
- envrionmental/species variables symbolised on biplots and maps

# SET UP
### Libraries
```{r}
library(plotly)
library(readxl)
```


### MAKE INPUTS

#### make full dataset to explore 
should include all env Vars and otus to explore, and make species richness variable

```{r}
#add species data and sp richness variable
env_sub_meta1<-left_join(env,otu_6,by="SampID")
env_sub_meta1$spRich<-rowSums(otu_6[,-c(1:which(colnames(otu_6)=="Tubularia_sp."))]!=0)

#rename X
env_sub_meta1$X <- env$X.y
env_sub_meta1 <-env_sub_meta1 %>% select (-c(X.y))

#add samplID
#env_sub_meta1$SampID<-envSel$SampID

```


### Add provisional biotope data to env file
Note that biotopes were last assigned in march 2022 and therefore there are some addional samples that have not yet got a biotope assigned. These should just be NAs

```{r}
biotopeInfo<-read_xlsx(file.path(dataPath, "inputs/MAREANO_provisional_biotope_classification_0322.xlsx"), sheet=1) %>%
  select(-c(x_coordinate_UTM33N, y_coordinate_UTM33N))

biotopeLookup<-read_xlsx(file.path(dataPath, "inputs/MAREANO_provisional_biotope_classification_0322.xlsx"), sheet=2)

env_sub_meta<-left_join(env_sub_meta1,biotopeInfo)
```


### a colour palette for discrete plots (see later in script)
```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# discrete colur pallete where you decide the number of colours - max is 74.
n <- 61
qual_col_pals <-brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector <-unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

pie(rep(1,n), col=sample(col_vector, n))
```




## MAKE ADDITIONAL VARS

**INSER YOURS HERE - E.G. COMBINED SPECIES LISTS?**

### seddan classes
```{r}
env_sub_meta<- env_sub_meta %>% 
  mutate(
    seddanName = case_when(
            seddan == "10" ~ "Morraine_indet", ########### ALL NEW NAMES NEED ADDING!
            seddan == "11" ~ "Morraine_thick",
            seddan == "15" ~ "Morraine_marginal",
            seddan == "140" ~ "Rock",
            seddan == "150" ~ "Sedimentary_rock",
            seddan == "200" ~ "Suspension",
            seddan == "201" ~ "Bottom_current",
            seddan == "202" ~ "Glaciomarine",
            seddan == "211" ~ "Contourite",
            seddan == "213" ~ "Debris_flow",
            seddan == "215" ~ "Channel_deposit",
            seddan == "217" ~ "Bioclastic",
            seddan == "219" ~ "Layered_over_debris_flow",
            seddan == "220" ~ "Carbonate_crust",
            seddan == "230" ~ "Morraine_glaciomarineCover",
            seddan == "240" ~ "Slide_w_cover",
            seddan == "241" ~ "Slide_part_cover"
    ))
```

### make xy interaction (helpful for visulisation in gnmds)

```{r}
env_sub_meta<- env_sub_meta %>% 
  mutate(
    XY = X*Y)
```


### List variables you can play with
```{r}
colnames(env_sub_meta)
```


# EXPLORATORY PLOTS
GIVEN HERE ARE EXAMPLES OF DIFFERENT TYPES OF PLOTS - you should modify the varaibles you are plotting at will, and customise this notebook for the split you are exploring.


**INSERT YOURS HERE - SEE BELOW FOR TEMPLATE CHUNKS**

#### X v Y - gnmds
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(env_sub_meta$gnmds1), max(env_sub_meta$gnmds1)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Geography (X v Y) - coloured by gnmds1") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_gnmds1.png"), #amend
       device = "png",
       dpi=300 )
```
#### X v Y - gnmds axis 2
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = gnmds2),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(env_sub_meta$gnmds2), max(env_sub_meta$gnmds2)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Geography (X v Y) - coloured by gnmds2") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_gnmds2.png"), #amend
       device = "png",
       dpi=300 )
```




#### X v Y - dissox
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = BO22_dissoxmean_bdmean),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(294, max(env_sub_meta$BO22_dissoxmean_bdmean)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Geography (X v Y) - coloured by dissolved oxygen") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_dissoxmean_294thresh.png"), #amend
       device = "png",
       dpi=300 )
```

#### X v Y with MLDmean_bathy
```{r}
cat_xy_ax1 <- ggplot(data = env_sub_meta,
              aes(x = X,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = MLDmean_bathy), #change at will - categorical vars only
             size = 1) +
  scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
 # scale_colour_brewer(palette = "col_vector") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by MLDmean-bathycategories") #amend

#cat_xy_ax1 #plot
ggplotly(cat_xy_ax1) #interactive plot
```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__XvY_MLDmean_bahty.png"), #amend
       device = "png",
       dpi=300 )

```
#### X v Y - chlomin
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = BO22_chloltmin_ss),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(env_sub_meta$BO22_chloltmin_ss), 0.06), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Geography (X v Y) - coloured by min long term chlorophyll") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_chloltmin_0pt06thresh.png"), #amend
       device = "png",
       dpi=300 )
```
#### X v Y - Tmean
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(env_sub_meta$Tmean_Robinson), max(env_sub_meta$Tmean_Robinson)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Geography (X v Y) - coloured by Temperature (meanR)") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_Tean_Rob.png"), #amend
       device = "png",
       dpi=300 )
```
#### X v Y - Smax Roinson
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Smax_Robinson),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(env_sub_meta$Smax_Robinson), max(env_sub_meta$Smax_Robinson)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Geography (X v Y) - coloured by Salinity (maxR)") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_Smax_Rob.png"), #amend
       device = "png",
       dpi=300 )
```


#### X v Y - Smax blended
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = salt_max),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(env_sub_meta$salt_max), max(env_sub_meta$salt_max)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Geography (X v Y) - coloured by Salinity (max BlendedModel)") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_Smax.png"), #amend
       device = "png",
       dpi=300 )
```
#### X v Y with seddanName
```{r}
cat_xy_ax1 <- ggplot(data = env_sub_meta,
              aes(x = X,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = seddanName), #change at will - categorical vars only
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by sediment origin") #amend

#cat_xy_ax1 #plot
ggplotly(cat_xy_ax1) #interactive plot
```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__XvY_seddanName.png"), #amend
       device = "png",
       dpi=300 )

```
#### GNMDS BIPLOT - X
```{r}
gnmds_spR <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by X/latitude")+ #amend
  geom_point(aes(colour = X)) + #change at will - continuous vars only
  scale_colour_gradientn(limits = c(min(env_sub_meta$X),max(env_sub_meta$X)),#match variable acting as colour
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

#p_spR #plot
ggplotly(gnmds_spR) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030_gnmds_Xlat.png"), #amend
       device = "png",
       dpi=300 )
```

#### GNMDS BIPLOT - Y
```{r}
gnmds_spR <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by Y/longitude")+ #amend
  geom_point(aes(colour = Y)) + #change at will - continuous vars only
  scale_colour_gradientn(limits = c(min(env_sub_meta$Y),max(env_sub_meta$Y)),#match variable acting as colour
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

#p_spR #plot
ggplotly(gnmds_spR) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030_gnmds_Ylon.png"), #amend
       device = "png",
       dpi=300 )
```
#### GNMDS BIPLOT - X*Y
```{r}
gnmds_spR <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by XYinteraction")+ #amend
  geom_point(aes(colour = XY)) + #change at will - continuous vars only
  scale_colour_gradientn(limits = c(min(env_sub_meta$XY),max(env_sub_meta$XY)),#match variable acting as colour
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

#p_spR #plot
ggplotly(gnmds_spR) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030_gnmds_XYinteraction.png"), #amend
       device = "png",
       dpi=300 )
```


#### X v Y - XY for colour comparison in gnmds
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = XY),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(env_sub_meta$XY), max(env_sub_meta$XY)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Geography (X v Y) - coloured by X*Y interaction (for comparison vals/colours)") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_XYinteraction.png"), #amend
       device = "png",
       dpi=300 )
```

-----------------



# TEMPLATE CHUNKS TO COPY FROM

## Plots with CONTINUOUS colour scales (variables)

### VARIABLE BIPLOT
example -
#### Bathy v Temp w gnmds ax 1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson, #change at will
                          y = bathy)) +#change at will
  theme_classic() +
  geom_point(aes(colour = gnmds1),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(gnmds1),max(gnmds1)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by gnmds r6 ax1") #amend

#tb_ax1 #plot
ggplotly(tb_ax1) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030_bathy_v_temp_nmds1.png"), #amend
       device = "png",
       dpi=300 )
```

#### GNMDS BIPLOT
```{r}
gnmds_spR <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by species Richness")+ #amend
  geom_point(aes(colour = spRich)) + #change at will - continuous vars only
  scale_colour_gradientn(limits = c(min(spRich),max(spRich)),#match variable acting as colour
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

#p_spR #plot
ggplotly(gnmds_spR) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030_bathy_gnmds_spR.png"), #amend
       device = "png",
       dpi=300 )
```

#### DCA BIPLOT
```{r}
dca_spR <- ggplot(data = env_sub_meta,
                     aes(x = dca1,
                         y = dca2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("DCA coloured by species Richness")+ #amend
  geom_point(aes(colour = spRich)) + #change at will - continuous vars only
  scale_colour_gradientn(limits = c(min(spRich),max(spRich)),#match variable acting as colour
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

#p_spR #plot
ggplotly(dca_spR) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030_bathy_dca_spR.png"), #amend
       device = "png",
       dpi=300 )
```


### MAP OF VARIABLE
example -
#### X v Y - temp
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(Tmean_Robinson), max(Tmean_Robinson)), #match variable acting as colour
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__X_v_Y_tmeanR.png"), #amend
       device = "png",
       dpi=300 )
```




## Plots with DISCRETE colour scales (variables)

### MAKE VARIABLE CATEGORICAL
#### categorise gnmds
```{r}
env_sub_meta$ax1cat<-cut(env_sub_meta$gnmds1, 
      breaks=c(-3.2,-3,-2,-1,0,1,2,3,3.46))# need min val first and max val last, then you decide intermediate
env_sub_meta$ax2cat<-cut(env_sub_meta$gnmds2, 
      breaks=c(-1.9,-1,0,1,2,3,4,4.9))# need min val first and max val last, then you decide intermediate

#same method, bu add custom labels
env_sub_meta$dissoxav305<-cut(env_sub_meta$BO22_dissoxmean_bdmean, 
     breaks=c(282.5,305,372.2),# need min val first and max val last, then you decide intermediate 
     labels=c("lowO2","hiO2")) #custom labels which range from min to break,...,then break to max
```

### VARIABLE BIPLOTS
#### Gnmds w gmorph
```{r}

gnmds_gmo <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by landscape class",
          subtitle = "First run") +
  geom_point(aes(colour = factor(gmorph))) + #change at will - categorical vars only
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

gnmds_gmo #plot
ggploty(gnmds_gmo)

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__gnmds_gmo.png"), #amend
       device = "png",
       dpi=300 )

```

### MAP OF CATEGORICAL VARIABLE
#### X v Y with gnmds ax 1 as HC categories
```{r}
cat_xy_ax1 <- ggplot(data = env_sub_meta,
              aes(x = X,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = ax1cat), #change at will - categorical vars only
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by gnmds axis 1 HC units") #amend

#cat_xy_ax1 #plot
ggplotly(cat_xy_ax1) #interactive plot
```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__XvY_ax1cat.png"), #amend
       device = "png",
       dpi=300 )

```


## GNMDS with biotope
```{r}
p_gmo <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by biotope")+
  geom_point(aes(colour = factor(env_sub_meta$BiotopeDesc))) +
  scale_colour_manual(values=col_vector) +# non-ordered colourblind pallette
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour="none", size="none")

ggplotly(p_gmo)
```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__gnmds_biotope.png"), #amend
       device = "png",
       dpi=300 )

```
## map with biotope
```{r}
p_gmo <- ggplot(data = env_sub_meta,
                     aes(x = X,
                         y = Y)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("Map coloured by biotope")+
  geom_point(aes(colour = factor(env_sub_meta$BiotopeDesc))) +
  scale_colour_manual(values=col_vector) # non-ordered colourblind pallette
 

ggplotly(p_gmo)
```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__map_biotope.png"), #amend
       device = "png",
       dpi=300 )

```

# COMBINE SEVERAL PLOTS
```{r}
comp<-xy_t+cat_y_ax1+p_spR
```
##### Save the plot
```{r}
##### Save some outputs
ggexport(comp,
          filename = file.path(dataPath,"outputs/SHELF_SOFT_1030__compare.png"), #amend
          width = 1500, #amend with/height for expected shape ratio
          height = 500)

```



# Save environment

EDIT AREA FIRST!

```{r}
#save.image("I:/Scripts/deepseaNiN/Renv_deepseaNiN_SHELF_SOFT_1030_.RData") # edit area first

```

NOW CLEAR ENVIRONMENT BEFORE RUNNING FOR NEW AREA

