---
title: "MAREANO - NiN:splitting biogeographic regions_SHELF oct 1 sample per R station"
output: html_notebook
---

Asssumes you have aready run:
- config
- workflow_deepseaNiN_Start

You now have the objects needed to explore the dataset and find the best way to apply biogeographic splits. This is a pre-step before running the workflow again on the split dataset to explore possible LKMs.

# Finding Regional (biogeographic) Splits

This requires exploring the dataset to theorise the best ways to split the data with:
- environmental biplots/colourings


then applying the splits. 


### Libraries
```{r}
library(plotly)
library(readxl)
```




## Exploring biplots
#### make full dataset to explore 
should include all env Vars and otus to explore, and make species richness variable

```{r}
#add species data and sp richness variable
env_sub_meta1<-left_join(env,otu_6, by="SampID")
env_sub_meta1$spRich<-rowSums(otu_6[,-c(1:which(colnames(otu_6)=="Zoanthidae"))]!=0)

#rename X
env_sub_meta1$X <- env$X.y
#env_sub_meta1 <-env_sub_meta1 %>% select (-c(X.y))

#add samplID
#env_sub_meta1$SampID<-envSel$SampID

```


### Add provisional biotope data to env file
Note that biotopes were last assigned in march 2022 and therefore there are some addional samples that have not yet got a biotope assigned. These should just be NAs
```{r}
biotopeInfo<-read_xlsx(file.path(dataPath, "inputs/MAREANO_provisional_biotope_classification_0322.xlsx"), sheet=1) %>%
  select(-c(x_coordinate_UTM33N, y_coordinate_UTM33N))

biotopeLookup<-read_xlsx(file.path(dataPath, "inputs/MAREANO_provisional_biotope_classification_0322.xlsx"), sheet=2)

env_sub_meta<-left_join(env_sub_meta1,biotopeInfo, by="SampID")
```


### a colour palette for discrete plots (see later in script)

```{r}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

# discrete colur pallete where you decide the number of colours - max is 74.
n <- 61
qual_col_pals <-brewer.pal.info[brewer.pal.info$category == 'qual',]
col_vector <-unlist(mapply(brewer.pal, qual_col_pals$maxcolors, rownames(qual_col_pals)))

pie(rep(1,n), col=sample(col_vector, n))
```

### List variables you can play with
```{r}
colnames(env_sub_meta)
```
# MAKE MORE VARS

```{r}

```


# EXPLORATORY PLOTS
GIVEN HERE ARE EXAMPLES OF DIFFERENT TYPES OF PLOTS - you should modify the varaibles you are plotting at will, and customise this notebook for the split you are exploring.


**INSERT YOURS HERE - SEE BELOW FOR TEMPLATE CHUNKS**

## GNMDS with PoriferaBig
```{r}
gnmds_porBig<- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by Porifera Big",
          subtitle = "First run") +
  geom_point(aes(colour = env_sub_meta$Porifera_big)) +
  scale_colour_gradientn(limits = c(0.001,max(env_sub_meta$Porifera_big)),
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

ggplotly(gnmds_porBig)
```

## GNMDS with Lophelia
```{r}
gnmds_Loph<- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by Lophelia pertusa",
          subtitle = "First run") +
  geom_point(aes(colour = env_sub_meta$Lophelia_pertusa)) +
  scale_colour_gradientn(limits = c(0.001,max(env_sub_meta$Lophelia_pertusa)),
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

ggplotly(gnmds_Loph)
```


## GNMDS with biotope
```{r}
p_gmo <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by biotope")+
  geom_point(aes(colour = factor(env_sub_meta$BiotopeDesc))) +
  scale_colour_manual(values=col_vector) +# non-ordered colourblind pallette
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour="none", size="none")

ggplotly(p_gmo)
```

## GNMDS with seddan

### seddan classes
```{r}
env_sub_meta<- env_sub_meta %>% 
  mutate(
    seddanName = case_when(
            seddan == "10" ~ "Morraine_indet", ########### ALL NEW NAMES NEED ADDING!
            seddan == "11" ~ "Morraine_thick",
            seddan == "15" ~ "Morraine_marginal",
            seddan == "140" ~ "Rock",
            seddan == "150" ~ "Sedimentary_rock",
            seddan == "200" ~ "Suspension",
            seddan == "201" ~ "Bottom_current",
            seddan == "202" ~ "Glaciomarine",
            seddan == "211" ~ "Contourite",
            seddan == "213" ~ "Debris_flow",
            seddan == "215" ~ "Channel_deposit",
            seddan == "217" ~ "Bioclastic",
            seddan == "219" ~ "Layered_over_debris_flow",
            seddan == "220" ~ "Carbonate_crust",
            seddan == "230" ~ "Morraine_glaciomarineCover",
            seddan == "240" ~ "Slide_w_cover",
            seddan == "241" ~ "Slide_part_cover"
    ))
```


```{r}
p_gmo <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by sediment origin")+
  geom_point(aes(colour = factor(env_sub_meta$seddan))) +
  scale_colour_manual(values=col_vector) +# non-ordered colourblind pallette
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour="none", size="none")

ggplotly(p_gmo)
```



# TEMPLATE CHUNKS TO COPY FROM

## Plots with CONTINUOUS colour scales (variables)

### VARIABLE BIPLOT
example -
#### Bathy v Temp w gnmds ax 1
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = Tmean_Robinson, #change at will
                          y = bathy)) +#change at will
  theme_classic() +
  geom_point(aes(colour = gnmds1),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(gnmds1),max(gnmds1)), #match variable acting as colour
                        colors=c('red','yellow','green'))+
  ggtitle("Temp (mean Robinson) vs Bathy - coloured by gnmds r6 ax1") #amend

#tb_ax1 #plot
ggplotly(tb_ax1) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELFoct1perR_bathy_v_temp_nmds1.png"), #amend
       device = "png",
       dpi=300 )
```

#### GNMDS BIPLOT
```{r}
gnmds_spR <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by species Richness")+ #amend
  geom_point(aes(colour = spRich)) + #change at will - continuous vars only
  scale_colour_gradientn(limits = c(min(spRich),max(spRich)),#match variable acting as colour
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

#p_spR #plot
ggplotly(gnmds_spR) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELFoct1perR_bathy_gnmds_spR.png"), #amend
       device = "png",
       dpi=300 )
```

#### DCA BIPLOT
```{r}
dca_spR <- ggplot(data = env_sub_meta,
                     aes(x = dca1,
                         y = dca2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("DCA coloured by species Richness")+ #amend
  geom_point(aes(colour = spRich)) + #change at will - continuous vars only
  scale_colour_gradientn(limits = c(min(spRich),max(spRich)),#match variable acting as colour
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

#p_spR #plot
ggplotly(dca_spR) #interactive plot
```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELFoct1perR_bathy_dca_spR.png"), #amend
       device = "png",
       dpi=300 )
```


### MAP OF VARIABLE
example -
#### X v Y - temp
```{r}
xy_t<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),#change at will - continuous vars only
             size = 1) +
  scale_colour_gradientn(limits = c(min(Tmean_Robinson), max(Tmean_Robinson)), #match variable acting as colour
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp") #amend

#xy_t #plot
ggplotly(xy_t) #interactive plot

```
##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELFoct1perR__X_v_Y_tmeanR.png"), #amend
       device = "png",
       dpi=300 )
```




## Plots with DISCRETE colour scales (variables)

### MAKE VARIABLE CATEGORICAL
#### categorise gnmds
```{r}
env_sub_meta$ax1cat<-cut(env_sub_meta$gnmds1, 
      breaks=c(-3.2,-3,-2,-1,0,1,2,3,3.46))# need min val first and max val last, then you decide intermediate
env_sub_meta$ax2cat<-cut(env_sub_meta$gnmds2, 
      breaks=c(-1.9,-1,0,1,2,3,4,4.9))# need min val first and max val last, then you decide intermediate

#same method, bu add custom labels
env_sub_meta$dissoxav305<-cut(env_sub_meta$BO22_dissoxmean_bdmean, 
     breaks=c(282.5,305,372.2),# need min val first and max val last, then you decide intermediate 
     labels=c("lowO2","hiO2")) #custom labels which range from min to break,...,then break to max
```

### VARIABLE BIPLOTS
#### Gnmds w gmorph
```{r}

gnmds_gmo <- ggplot(data = env_sub_meta,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by landscape class",
          subtitle = "First run") +
  geom_point(aes(colour = factor(gmorph))) + #change at will - categorical vars only
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

gnmds_gmo #plot
ggploty(gnmds_gmo)

```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELFoct1perR__gnmds_gmo.png"), #amend
       device = "png",
       dpi=300 )

```

### MAP OF CATEGORICAL VARIABLE
#### X v Y with gnmds ax 1 as HC categories
```{r}
cat_xy_ax1 <- ggplot(data = env_sub_meta,
              aes(x = X.y,
                  y = Y)) +
  theme_classic() +
  geom_point(aes(colour = ax1cat), #change at will - categorical vars only
             size = 1) +
 # scale_colour_manual(values=cbPalette) +# non-ordered colourblind pallette
  scale_colour_brewer(palette = "Spectral") + # ordered colourblind pallette
  ggtitle("Easting vs Northing - coloured by gnmds axis 1 HC units") #amend

#cat_xy_ax1 #plot
ggplotly(cat_xy_ax1) #interactive plot
```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/SHELFoct1perR__XvY_ax1cat.png"), #amend
       device = "png",
       dpi=300 )

```


# COMBINE SEVERAL PLOTS
```{r}
comp<-xy_t+cat_y_ax1+p_spR
```
##### Save the plot
```{r}
##### Save some outputs
ggexport(comp,
          filename = file.path(dataPath,"outputs/SHELFoct1perR__compare.png"), #amend
          width = 1500, #amend with/height for expected shape ratio
          height = 500)

```



# Save environment

EDIT AREA FIRST!

```{r}
#save.image("I:/Scripts/deepseaNiN/Renv_deepseaNiN_SHELFoct1perR_.RData") # edit area first

```

NOW CLEAR ENVIRONMENT BEFORE RUNNING FOR NEW AREA

