---
title: "MAREANO - NiN - sandwave "
authors: Thijs van Son, Rune Halvorsen, Rebecca Ross, Genoveva Gonzales-Mirelis, Margaret Dolan
date: "Last Rendered on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 2
    toc_float: yes
    fig_width: 7
    fig_height: 7
always_allow_html: true 
---

## Packages
```{r}
library(magrittr)
library(dplyr)
library(ggcorrplot)
library(ggpubr)
library(patchwork)
# library("here")
# library("bookdown")
# library("downloadthis")
library(vegan)
library(plyr)
library(e1071)
library(tidyverse)
library(viridis)
library(GGally)
library(ggrepel)
library(ggordiplots)
library(readr)
library(RColorBrewer)
library(oce) 
library(plotly)
library(purrr)
library(furrr)


source("gg_ordisurf_viridis.R")
```

<!-- LOAD OBJECTS IF HAVE RUN ALREADY -->
<!-- ```{r} -->
<!-- geodist_pa<-readRDS(file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_geodist_pa.rds")) -->
<!-- geodist_r6<-readRDS(file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_geodist_r6.rds")) -->

<!-- mds_pa<-readRDS(file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_mds_pa.rds")) -->
<!-- mds_r6<-readRDS(file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_mds_r6.rds")) -->
<!-- #mds_r6_1000<-readRDS(file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_mds_r6_1000rep.rds")) -->

<!-- ep<-0.8 #epsilon (to avoid having to find and run just this line from code blocks where the mds objects were made) -->
<!-- ``` -->


## Data

for the sandwave case study we are not using a split - input the original full dataset and select the relevant area later


```{r}

env_sort <- read.csv(file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_LDens_env_sort_2022-10-14.csv")) %>% as.data.frame
otu_sort <- read.csv(file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_LDens_otu_sort_2022-10-14.csv")) %>% as.data.frame

table(env_sort$SampID2 == otu_sort$SampID)


#must be same length and equal to unique sample number length - the ordered lists are assumed to be directly relatable on a row by row basis

dim(env_sort)
dim(otu_sort)
```


### Additional env data

You can look up EUNIS column meanings here - p14 on: https://archimer.ifremer.fr/doc/00723/83528/88577.pdf
Sediment miljø (erosion/deposition stuff) https://www.ngu.no/upload/Kartkatalog/Presentasjonsregler_Marin_Sedimentasjonsmiljo.pdf
Sediment dannelse (genesis) https://www.ngu.no/upload/Kartkatalog/Presentasjonsregler_Marin_SedimentDannelse.pdf
```{r}
envExtra<-read.csv(file.path(dataPath,"inputs/samples_env_plus.csv")) %>% as.data.frame
envExtra<-envExtra %>% select(c("SampID2", 
                                #"Energy",#EUNIS #TEXT
                                #"Biozone",#EUNIS #TEXT
                                #"Substrate",#EUNIS #TEXT
                                #"EUNIScomb",#EUNIS #TEXT
                                #"EUNIScombD",#EUNIS #TEXT
                                #"Allcomb",#EUNIS #TEXT
                                #"AllcombD",#EUNIS #TEXT
                                #"MSFD_BBHT",#EUNIS #TEXT
                                #"EUNIS2019C",#EUNIS #TEXT
                                #"EUNIS2019D",#EUNIS #TEXT
                                #"All2019D",#EUNIS #TEXT
                                #"All2019DL2",#EUNIS #TEXT
                                "sedmil", #erosion/deposition #CATEGORICAL
                                "seddan", #sed genesis #CATEGORICAL 
                                "sw_spd_max", #sandwave model max current speed
                                "sw_spd_mean", #sandwave model mean current speed
                                "TOC0.10cm_median" #Total Organic Carbon (Markus Diesing preliminary model 10/2022 top 10cm prediction)
))

env_sort<-left_join(env_sort,envExtra, by=c("SampID"="SampID2"))
```



#### remove any variables with not enough coverage
```{r}

env_sort <- env_sort %>% select(-c("BO22_lightbotmean_bdmean",
                               "BO22_lightbotltmax_bdmean",
                               "BO22_lightbotltmin_bdmean",
                               "BO22_lightbotrange_bdmean"))
```


#### Plot map to check location of samples
```{r}
env_sort_locations<- ggplot(data = env_sort,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Location of samples in sorted env file")

env_sort_locations
```

<!-- #### Plot map to check SampleIDs -->
<!-- ```{r} -->
<!-- env_sort_SampID<- ggplot(data = env_sort, -->
<!--                       aes(x = X.y, -->
<!--                           y = Y)) + -->
<!--   theme_classic() + -->
<!--   geom_point(aes(colour = factor(SampID)), size=1) + -->
<!--   ggtitle("Location of samplesIDs in sorted env file") -->

<!-- ggplotly(env_sort_SampID) -->

<!-- ``` -->



### Data cleaning
by using complete cases here since env_sort contains the sandwave current data, we automatically select the sandwave area
```{r}
## Removing NAs
otuCompl <- otu_sort[complete.cases(env_sort[, -c(1:which(colnames(env_sort)=="bathy")-1)]), ]
envCompl <- env_sort[complete.cases(env_sort[, -c(1:which(colnames(env_sort)=="bathy")-1)]), ]

## Removing observations with less than 4 OTUs

otuCompl$SpR<-rowSums(otuCompl[,c(3:(which(colnames(otuCompl)=="Zoanthidae")))]!=0)

sel <- otuCompl$SpR>4
inv.sel<-otuCompl$SpR<4


otuCompl<-otuCompl %>% select(-c("SpR"))

otuSel <- otuCompl[sel, ]
envSel <- envCompl[sel, ]


## Removing phosphate
#envSel <- envSel %>% select(-phosphate_mean.tif)


dim(otuSel); dim(envSel)
```
#### Show what samples are left after complete cases and >4 OTUs filters
```{r}
env_sel_locations<- ggplot(data = envSel,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Location of samples in sorted env file")

env_sel_locations
```

#### Show what got removed in complete cases filter
```{r}
envCompl_ccrem<-env_sort%>%filter(!SampID%in%envCompl$SampID)

envCompl_ccrem_locations<- ggplot(data = envCompl_ccrem,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Location of removed complete case samples resulting in envSel file")

envCompl_ccrem_locations
```
#### Show what got removed in <4 OTUs filter


```{r}

env.invSel <- envCompl[inv.sel, ]

env.invSel_locations<- ggplot(data = env.invSel,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = bathy),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Location of removed samples with <4 OTUs resulting in envSel file")

env.invSel_locations

```



#### Plot map to check SampleIDs
```{r}
env_sort_SampID<- ggplot(data = envSel,
                      aes(x = X.y,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = factor(SampID)), size=1) +
  ggtitle("Location of samplesIDs in sorted env file")

ggplotly(env_sort_SampID)

```




### Whole dataset for first run
```{r}
otu<-otuSel[,-c(1:which(colnames(otuSel)=="Acesta_excavata")-1)]
env<-envSel[, -c(1:which(colnames(env_sort)=="bathy")-1)]

table(is.na(otu))

table(is.na(env))

#str(otuSel)
#str(envSel)
```



## Abundance weighting
Make function
You might have to drop variables that have been imported as character
```{r}

otu_pa <- decostand(x = otu,
                    method = "pa")

# y = ax^w         # power transformation formula
dt <- otu       # species data to transform
x_mn <- min(dt[dt > 0])
x_mx <- max(dt)
rng <- 6           # abundance range
w <- log(rng) / (log(x_mx) - log(x_mn))
a <- x_mn^(-w)
# otu_6 <- a * dt[, -c(1:3)]^w
otu_6 <- a * dt^w
range(otu_6)
```

## Ordination methods
### DCA PA
```{r}
Sys.time()
dca_pa <- decorana(veg = otu_pa)

print(dca_pa, head=T)
```


### GNMDS PA

#### Distances - don't run if loading saved object

```{r}
## Bray-Curtis
dist_pa <- vegdist(x = otu_pa, method = "bray")

## Geodist
ep <- 0.8     # epsilon
geodist_pa <- isomapdist(dist = dist_pa, epsilon = ep)
```


##### Save the result - don't run if loading saved object
```{r}
saveRDS(geodist_pa,
        file = (file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_Ldens_geodist_pa.rds")))
```

#### Ordination

#### monoMDS - don't run if loading saved object
took 6.4mins with jonatan's paralell solution (usually ~10mins)

Using old version without furrr
```{r}
# UNCOMMENT IN THIS CHUNK IF YOU CANNOT RUN THE PARALLEL SOLUTION
#
# Sys.time()
# d <- 2
# mds_pa <- list()
# 
# for (i in 1:100) {
#   mds_pa[[i]]<-monoMDS(geodist_pa,
#                     matrix(c(runif(dim(otu_pa)[1]*d)),
#                            nrow = dim(otu_pa)[1]),
#                     k = d,
#                     model = "global",
#                     maxit = 2000,
#                     smin = 1e-7,
#                     sfgrmin = 1e-7)
# }
# Sys.time()



## COMMENT OUT FROM HERE IN THIS CHUNK IF YOU CANNOT RUN THE PARALLEL SOLUTION
plan(multisession)

d <- 2

i <-200 # number of reps

List_geodist_pa <- lapply(seq_len(i), function(X) geodist_pa) # makes a list with the input data repeated as many times as reps wanted
start_t<-Sys.time()
mds_pa<-furrr::future_map(List_geodist_pa, 
                  function(x) monoMDS(x,
                                      matrix(c(runif(dim(otu_6)[1]*d)),
                                             nrow = dim(otu_6)[1]),
                                      k = d,
                                      model = "global",
                                      maxit = 2000,
                                      smin = 1e-7,
                                      sfgrmin = 1e-7),
                  .progress = TRUE)
Sys.time() - start_t

```

##### Save the result - don't run if loading saved object
can take a few mins
```{r}
saveRDS(mds_pa,
        file = file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_Ldens_mds_pa.rds"))
```


##### Best nmds solution - PA


```{r}
# Loading mds results
# mds <- readRDS(file = "../SplitRev2_mds_pa_full.rds")

## Extracting the stress of each nmds iteration
mds_stress_pa<-unlist(lapply(mds_pa, function(v){v[[22]]})) 

ordered_pa <-order(mds_stress_pa)

## Best, second best, and worst solution
mds_stress_pa[ordered_pa[1]]
mds_stress_pa[ordered_pa[2]]
mds_stress_pa[ordered_pa[10]]

## Scaling of axes to half change units and varimax rotation by postMDS
mds_best_pa<-postMDS(mds_pa[[ordered_pa[1]]],
                  geodist_pa, 
                  pc = TRUE, 
                  halfchange = TRUE, 
                  threshold = ep)     # Is this threshold related to the epsilon above?
mds_best_pa

mds_secbest_pa <- postMDS(mds_pa[[ordered_pa[2]]],
                          geodist_pa, 
                          pc = TRUE, 
                          halfchange = TRUE, 
                          threshold = ep)
mds_secbest_pa

## Procrustes comparisons
procr_pa <- procrustes(mds_best_pa,
                    mds_secbest_pa,
                    permutations=999)
protest(mds_best_pa,
        mds_secbest_pa,
        permutations=999)

plot(procr_pa)

png(file=file.path(dataPath,"outputs/sandwave_Ldens_procrustes_pa.png"), width=1000, height=700)
plot(procr_pa)
dev.off()
```

##### Correlation of axis: DCA vs NMDS - PA
```{r}
# Extracting ordination axis
ax <- 2
axis_pa <- cbind(mds_best_pa$points,
                 scores(dca_pa,
                        display = "sites",
                        origin = TRUE)[, 1:ax])

ggcorr(axis_pa, 
       method=c("everything","kendall"), 
       label = TRUE,
       label_size = 3, 
       label_color = "black",  
       nbreaks = 8,
       label_round = 3,
       low = "red",
       mid = "white",
       high = "green")


```

##### Save the figure
```{r}
ggsave(filename = file.path(dataPath,"outputs/sandwave_Ldens_correlationPCAvsNMDS_PA.png"),
       device = "png",
       dpi=300 )

# Switching direction of NMDS1
# mds_best$points[, 1] <- -mds_best$points[, 1]
```

### DCA R6
```{r}
dca_r6 <- decorana(veg = otu_6)

print(dca_r6, head=T)
```


### GNMDS R6
#### Distances - don't run if loading saved object
```{r}
## Bray-Curtis
dist_r6 <- vegdist(x = otu_6, method = "bray")

## Geodist
ep <- 0.80     # epsilon
geodist_r6 <- isomapdist(dist = dist_r6, epsilon = ep)
```

##### Save the result - don't run if loading saved object
```{r}
saveRDS(geodist_r6,
        file = (file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_Ldens_geodist_r6.rds")))
```


#### Ordination 

##### 200 reps - don't run if loading saved object
(Jonatan's parallel solution took 5.4mins, before it took 10mins)

Using older version without furrr.
```{r}

## UNCOMMENT IN THIS CHUNK IF YOU CANNOT RUN THE PARALLEL SOLUTION

# monoMDS
# d <- 2
# mds_r6 <- list()
# 
# Sys.time()
# for (i in 1:200) {
#   mds_r6[[i]]<-monoMDS(geodist_r6,
#                     matrix(c(runif(dim(otu_6)[1]*d)),
#                            nrow = dim(otu_6)[1]),
#                     k = d,
#                     model = "global",
#                     maxit = 2000,
#                     smin = 1e-7,
#                     sfgrmin = 1e-7)
# }
# Sys.time()

## COMMENT OUT FROM HERE IN THIS CHUNK IF YOU CANNOT RUN THE PARALLEL SOLUTION

## monoMDS
plan(multisession)

d <- 2

i <-200 # number of reps

List_geodist_r6 <- lapply(seq_len(i), function(X) geodist_r6) # makes a list with the input data repeated as many times as reps wanted
start_t<-Sys.time()
mds_r6<-furrr::future_map(List_geodist_r6, 
                  function(x) monoMDS(x,
                                      matrix(c(runif(dim(otu_6)[1]*d)),
                                             nrow = dim(otu_6)[1]),
                                      k = d,
                                      model = "global",
                                      maxit = 2000,
                                      smin = 1e-7,
                                      sfgrmin = 1e-7),
                  .progress = TRUE)
Sys.time() - start_t

```

##### Save the result - don't run if loading saved object
```{r}
saveRDS(mds_r6,
        file = file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_Ldens_mds_r6.rds")) 
```

##### Best nmds solution r6 200 rep 
```{r, eval=TRUE}
# Loading geodist object
# geodist_nfi <- readRDS(file = "../SplitRev2_geodist_nfi.rds")

# Loading mds results
# mds <- readRDS(file = "../SplitRev2_mds.rds")

## Extracting the stress of each nmds iteration
mds_stress_r6<-unlist(lapply(mds_r6, function(v){v[[22]]})) 

ordered_r6 <-order(mds_stress_r6)

## Best, second best, and worst solution
mds_stress_r6[ordered_r6[1]]
mds_stress_r6[ordered_r6[2]]
mds_stress_r6[ordered_r6[100]]

## Scaling of axes to half change units and varimax rotation by postMDS
mds_best_r6<-postMDS(mds_r6[[ordered_r6[1]]],
                     geodist_r6, 
                     pc = TRUE, 
                     halfchange = TRUE, 
                     threshold = ep)     # Is this threshold related to the epsilon above?
mds_best_r6

mds_secbest_r6<-postMDS(mds_r6[[ordered_r6[2]]],
                        geodist_r6, 
                        pc = TRUE, 
                        halfchange = TRUE, 
                        threshold = ep)
mds_secbest_r6

## Procrustes comparisons
procr_r6 <- procrustes(mds_best_r6,
                       mds_secbest_r6,
                       permutations=999)
protest(mds_best_r6,
        mds_secbest_r6,
        permutations=999)

plot(procr_r6)

png(file.path(dataPath,"outputs/sandwave_Ldens_procrustes_r6.png"), width=1000, height=700,) #added 1000
plot(procr_r6)
dev.off()
```



##### Correlation of axis: DCA vs NMDS 
retain the 200 rep version
```{r}
# Extracting ordination axis
ax <- 2
axis_r6 <- cbind(mds_best_r6$points,
                 scores(dca_r6,
                        display = "sites",
                        origin = TRUE)[, 1:ax])

ggcorr(axis_r6, 
       method=c("everything","kendall"), 
       label = TRUE,
       label_size = 3, 
       label_color = "black",  
       nbreaks = 8,
       label_round = 3,
       low = "red",
       mid = "white",
       high = "green")

```

##### Save the figure
```{r}
ggsave(filename = file.path(dataPath,"outputs/sandwave_Ldens_correlationPCAvsNMDS_r6.png"),
       device = "png",
       dpi=300 )

# Switching direction of NMDS1
# mds_best$points[, 1] <- -mds_best$points[, 1]
```

### Plotting DCA & GNMDS
```{r}
## Adding scores to data frame
otu_6$gnmds1 <- mds_best_r6$points[, 1]
otu_6$gnmds2 <- mds_best_r6$points[, 2]
otu_6$dca1 <- scores(dca_r6, display = "sites", origin = TRUE)[, 1]
otu_6$dca2 <- scores(dca_r6, display = "sites", origin = TRUE)[, 2]

p_gnmds_r6 <- ggplot(data = otu_6,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS",
          subtitle = "First run") +
  geom_point(colour = "red") +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")

p_dca_r6 <- ggplot(data = otu_6,
                   aes(x = dca1,
                       y = dca2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("DCA",
          subtitle = "First run") +
  geom_point(colour = "red") +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")

p_gnmds_r6 + p_dca_r6


```

##### Save the figure
```{r}
ggsave(file.path(dataPath,"outputs/sandwave_Ldens_gnmds_dca_r6.png"),
       device = "png", 
       dpi=300)
```

## Check outlier species Richness
```{r}
otu_6$spRich<-rowSums(otu_6[,c(1:which(colnames(otu_6)=="Zoanthidae"))]!=0)

summary(otu_6$spRich)

p_spR <- ggplot(data = otu_6,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by species Richness",
          subtitle = "First run") +
  geom_point(aes(colour = spRich)) +
  scale_colour_gradientn(limits = c(min(otu_6$spRich),max(otu_6$spRich)),
                        colors=c('red','yellow','green'))+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")#+
  #guides(colour="none", size="none")

ggplotly(p_spR)
```



## Check outlier SampIDs
```{r}

otu_6$SampID<-envSel$SampID


p_gmo <- ggplot(data = otu_6,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by sample",
          subtitle = "First run") +
  geom_point(aes(colour = factor(SampID))) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour="none", size="none")

ggplotly(p_gmo)
```





------------------------

# Environmental Correlates 

If you are happy with proceeding (i.e. you have decent procrustes plots, and have identified and removed outliers with some understanding of why they are outliers) Then here you can continue with exploring environmental correlates from here.

NB it can be useful to proceed anyway if you want to try and explain why some samples are outliers, but if it is possible to deal with minor sets of outliers already then this is a good idea.




## Species-environment relationships
### Selecting ordination
```{r}
ord <- mds_best_r6

## Axis scores if selected ord is GNMDS
axis <- ord$points %>% as.data.frame

## Axis scores if selected ord is DCA
# axis <- scores(ord,
#                display = "sites",
#                origin = TRUE)[, 1:ax])

```

### Create additional variables
decided to make MLD-bathy vars
```{r}
env<-env %>% 
  mutate ("MLDmean_bathy"=MLDmean_Robinson-(bathy*-1),
          "MLDmin_bathy"=MLDmin_Robinson-(bathy*-1),
          "MLDmax_bathy"=MLDmax_Robinson-(bathy*-1))

env$MLDmean_bathy<-cut(env$MLDmean_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))
env$MLDmin_bathy<-cut(env$MLDmin_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))
env$MLDmax_bathy<-cut(env$MLDmax_bathy, 
      breaks=c(-2560, -20,20,130),#checked range of values first (min -2554, max 123)
      labels=c('belowMLD','onPycno','inMixLayer'))

env$swDensRob_avs<-swRho(salinity=env$Smean_Robinson,
                         temperature=env$Tmean_Robinson,
                         pressure=(env$bathy*-1),
                         eos="unesco")

```



### Correlation ordination axes and environmental variables
#### Removing non-env vars
```{r}
env_cont<-env%>% select(-c(landscape,sedclass,gmorph, MLDmean_bathy, MLDmax_bathy, MLDmin_bathy, #categorical
                           #optional, #not a var
                           MLDmax_Robinson, MLDmean_Robinson, MLDmin_Robinson,  #replaced by new vars
                           MLDsd_Robinson, #not meaningful
                           #Energy, Biozone, Substrate, EUNIScomb,EUNIScombD,Allcomb,AllcombD,MSFD_BBHT,EUNIS2019C,EUNIS2019D,All2019D,All2019DL2, #EUNIS TEXT
                           sedmil,seddan #sed erosion/genesis # NEW CATEGORICAL
                           ))
env_cont<-env_cont%>% mutate_if(is.integer,as.numeric)
env_corr <- env_cont #%>% select(-c(SampID))

# env_corr$coords.x1<-as.numeric(env_corr$coords.x1)
# env_corr$coords.x2<-as.numeric(env_corr$coords.x2)

env_corr[(!is.numeric(env_corr)),]
```

#### Correlations
```{r}
# Vector to hold correlations
cor_ax1 <- NULL
cor_ax2 <- NULL
pv_ax1 <- NULL
pv_ax2 <- NULL

# NMDS1
for( i in seq(length(env_corr))) {
  ct.i <- cor.test(axis$MDS1,
                   env_corr[, i],
                   method = "kendall")
  cor_ax1[i] <- ct.i$estimate
  pv_ax1[i] <- ct.i$p.value
}

# NMDS2
for( i in seq(length(env_corr))) {
  ct.i <- cor.test(axis$MDS2,
                   env_corr[, i],
                   method = "kendall")
  cor_ax2[i] <- ct.i$estimate
  pv_ax2[i] <- ct.i$p.value
}

cor_tab <- data.frame(env = names(env_corr),
                      ord_ax1 = cor_ax1,
                      pval_ax1 = pv_ax1,
                      ord_ax2 = cor_ax2,
                      pval_ax2 = pv_ax2)

cor_tab

write.csv(x = cor_tab,
          file = file.path(dataPath,"inputs/SPLITS/sandwave/sandwave_Ldens_cor-table_r6_200rep.csv"),
          row.names = FALSE)
```
### Dot chart to check for gaps in correlation

```{r}
cor_a1_sort<-cor_tab%>%
  mutate(abs_ord_ax1=abs(ord_ax1),
         abs_ord_ax2=abs(ord_ax2)) %>%
  arrange(desc(abs_ord_ax1))

cor_a2_sort<-cor_tab%>%
  mutate(abs_ord_ax1=abs(ord_ax1),
         abs_ord_ax2=abs(ord_ax2)) %>%
  arrange(desc(abs_ord_ax2))

dotchart(cor_a1_sort$abs_ord_ax1, main="Absolute (+/-) correlations between envVars and gnmds axis 1")

cor_cut<-0.4 #decide

cor_sel<-subset(cor_a1_sort,abs_ord_ax1>cor_cut)
cor_sel

as.data.frame(cor_sel$env)
```


### Sel env var (top corr)
```{r}
env_os <- env[, cor_sel$env]
env_os
str(env_os)


```


### Ordisurfs top corr
```{r}
ordsrfs <- list(length = ncol(env_os))

for (i in seq(ncol(env_os))) {
  os.i <- gg_ordisurf(ord = ord,
                      env.var = env_os[, i],
                      pt.size = 1,
                      # binwidth = 0.05,
                      var.label = names(env_os)[i],
                      gen.text.size = 10,
                      title.text.size = 15,
                      leg.text.size = 10
                      )
  
  ordsrfs[[i]] <- os.i$plot
}

ordsrfs_plt <- ggarrange(plotlist = ordsrfs,
                         nrow = 5,
                         ncol = 4)

ordsrfs_plt
```
##### Save some outputs
```{r}
ggexport(ordsrfs_plt,
          filename = file.path(dataPath,"outputs/sandwave_Ldens_ordisurfs_top_corr.png"),
          width = 2000,
          height = 2500)

```


### Sel env var (manual)
```{r}
env_os_m <- env[,c("Tmean_Robinson", #top corr
                  "salt_max", #top corr
                  "Smax_Robinson", #comparison to top corr
                  "swDensRob_avs", #top corr
                  "BO22_icecoverltmax_ss",#top corr ax2
                  "BO22_icecovermean_ss",#top corr
                  "BO22_dissoxmean_bdmean",#top corr
                  #"BO22_carbonphytoltmin_bdmean",#top corr - no clear gradient in ordisurf
                  "BO22_ppltmin_ss", #top corr
                  "X.y", #comparison to Y
                  "Y", #top corr
                  "spd_std", #top corr ax2 (blended model)
                  "CSpdsd_Robinson", #comparison to top corr ax2 (blended model)
                  "mud", #highest sed var ax1 + corr
                  "gravel",#highest sed var ax1 - corr
                  "BO22_silicateltmax_bdmean", #just under top corr ax1
                  "bathy" #intuitive for comparisons
                )]
env_os_m
str(env_os_m)

```


### Ordisurfs manually selected
```{r}
ordsrfs_m <- list(length = ncol(env_os_m))

for (i in seq(ncol(env_os_m))) {
  os.i_m <- gg_ordisurf(ord = ord,
                      env.var = env_os_m[, i],
                      pt.size = 1,
                      # binwidth = 0.05,
                      var.label = names(env_os_m)[i],
                      gen.text.size = 10,
                      title.text.size = 15,
                      leg.text.size = 10)
  
  ordsrfs_m[[i]] <- os.i_m$plot
}

ordsrfs_plt_m <- ggarrange(plotlist = ordsrfs_m,
                         nrow = 4,
                         ncol = 4)

ordsrfs_plt_m
```

##### Save some outputs
```{r}
ggexport(ordsrfs_plt_m,
          filename = file.path(dataPath,"outputs/sandwave_Ldens_ordisurfs_man_sel_domean.png"),
          width = 2000,
          height = 2000)

```




### Envfit
```{r}
## Select if any var should be excluded from envfit (makes less busy to read)
env_os_m_envfit<-env_os_m [,c("Tmean_Robinson", #top corr
                  "salt_max", #top corr
                  "Smax_Robinson", #comparison to top corr
                  "swDensRob_avs", #top corr
                  "BO22_icecoverltmax_ss",#top corr ax2
                  #"BO22_icecovermean_ss",#top corr
                  "BO22_dissoxmean_bdmean",#top corr
                  #"BO22_dissoxltmin_bdmean",#top corr
                  #"BO22_carbonphytoltmin_bdmean",#top corr - no clear gradient in ordisurf
                  "BO22_ppltmin_ss", #top corr
                  "X.y", #comparison to Y
                  "Y", #top corr
                  "spd_std", #top corr ax2 (blended model)
                 # "CSpdsd_Robinson", #comparison to top corr ax2 (blended model)
                  "mud", #highest sed var ax1 + corr
                  "gravel",#highest sed var ax1 - corr
                  "BO22_silicateltmax_bdmean", #just under top corr ax1
                  "bathy" #intuitive for comparisons
                
  
)]

colnames(env_os_m_envfit)<-c("T", #top corr
                  "sMx", #top corr
                  "SmaxR", #comparison to top corr
                  "swDensR", #top corr
                  "icecovmax",#top corr ax2
                  #"icecovav",#top corr
                  "dissoxav",#top corr
                  #"dissoxmin",#top corr
                  #"BO22_carbonphytoltmin_bdmean",#top corr - no clear gradient in ordisurf
                  "ppltmin", #top corr
                  "X", #comparison to Y
                  "Y", #top corr
                  "spd_std", #top corr ax2 (blended model)
                 # "CSsd", #comparison to top corr ax2 (blended model)
                  "mud", #highest sed var ax1 + corr
                  "gravel",#highest sed var ax1 - corr
                  "SiLtmax", #just under top corr ax1
                  "bathy" #intuitive for comparisons
                 )

## Envfot plot
gg_envfit(ord = ord,
          env = env_os_m_envfit,
          pt.size = 1)

## Envfit analysis

ef <- envfit(ord = ord,
             env = env_os_m_envfit,
             # na.rm = TRUE
             )
efDF <- as.data.frame(scores(ef,
                             display = "vectors"))
```

##### Save the plot
```{r}
ggsave(filename = file.path(dataPath,"outputs/sandwave_Ldens_EnvFit_man_sel_cln_domean.png"),
       device = "png",
       dpi=300 )
```
### Categorical envVar visualised on the mdsplots

#### use dataset with categorical var included

```{r}
env_vis<-env
env_vis$gnmds1 <- otu_6$gnmds1
env_vis$gnmds2 <- otu_6$gnmds2

env_vis$dca1 <- otu_6$dca1
env_vis$dca2 <- otu_6$dca2
```

#### gnmds w mld mean - bathy
```{r}
dca_int <- ggplot(data = env_vis,
                     aes(x = dca1,
                         y = dca2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("Interactive DCA sample ID plot",
          subtitle = "First run") +
  geom_point(aes(colour = factor(SampID))) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")

ggplotly(dca_int)
```





#### gnmds w mld mean - bathy
```{r}
p_mld <- ggplot(data = env_vis,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by proximity to mixed layer depth",
          subtitle = "First run") +
  geom_point(aes(colour = MLDmean_bathy)) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")

p_mld
```

#### gnmds w sedclass
##### deal with sedclass codes
```{r}
env_vis<- env_vis %>% 
  mutate(
    sedclassName = case_when(
            sedclass == "1" ~ "SedCoverR",
            sedclass == "5" ~ "Rock",
            sedclass == "20" ~ "Mud",
            sedclass == "21" ~ "MwBlock",
            sedclass == "40" ~ "sMud",
            sedclass == "80" ~ "mSand",
            sedclass == "100" ~ "Sand",
            sedclass == "110" ~ "gMud",
            sedclass == "115" ~ "gsMud",
            sedclass == "120" ~ "gmSand",
            sedclass == "130" ~ "gSand",
            sedclass == "150" ~ "MSG",
            sedclass == "160" ~ "sGravel",
            sedclass == "170" ~ "Gravel",
            sedclass == "175" ~ "GravBlock",
            sedclass == "185" ~ "SGBmix",
            sedclass == "205" ~ "S/MwB",
            sedclass == "206" ~ "S/MwG/B",
            sedclass == "215" ~ "SGBalt",
            sedclass == "300" ~ "HardSed",
            sedclass == "500" ~ "Biogenic"
                        
    )
  )
```

###### colour palette to cope with up to 25 categorical colours
```{r}
c25 <- c(
  "dodgerblue2", "#E31A1C", # red
  "green4",
  "#6A3D9A", # purple
  "#FF7F00", # orange
  "black", "gold1",
  "skyblue2", "#FB9A99", # lt pink
  "palegreen2",
  "#CAB2D6", # lt purple
  "#FDBF6F", # lt orange
  "gray70", "khaki2",
  "maroon", "orchid1", "pink1", "blue1", "steelblue4",
  "darkturquoise", "green1", "yellow4", "yellow3",
  "darkorange4", "brown"
)
```




```{r}

p_sed <- ggplot(data = env_vis,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by sediment class",
          subtitle = "First run") +
  geom_point(aes(colour = factor(sedclassName))) +
   scale_colour_manual(values=c25)+
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

p_sed
```

#### gnmds w lanscape
##### deal with sedclass codes
```{r}
env_vis<- env_vis %>% 
  mutate(
    landscapeName = case_when(
            landscape == "1" ~ "Strandflat",
            landscape == "21" ~ "ContSlope",
            landscape == "22" ~ "Canyon",
            landscape == "31" ~ "Valley",
            landscape == "32" ~ "Fjord",
            landscape == "41" ~ "DeepSeaPlane",
            landscape == "42" ~ "SlopePlain",
            landscape == "43" ~ "ShelfPlain",
            landscape == "431" ~ "shallowValley"
    )
  )
```


```{r}

p_land <- ggplot(data = env_vis,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by landscape class",
          subtitle = "First run") +
  geom_point(aes(colour = factor(landscapeName))) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

p_land
```


#### Gnmds w gmorph
```{r}

p_gmo <- ggplot(data = env_vis,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by landscape class",
          subtitle = "First run") +
  geom_point(aes(colour = factor(gmorph))) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour=guide_legend(ncol=2))

p_gmo
```


```{r}
cat_var_plots<-p_mld+p_sed+p_land+p_gmo
```
##### Save the plot
```{r}
ggexport(cat_var_plots,
          filename = file.path(dataPath,"outputs/sandwave_Ldens_gnmds_catvar.png"),
          width = 1000,
          height = 800)
```

#### Sample identification in the mdsplot

```{r}
p_gmo <- ggplot(data = env_vis,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS coloured by sample",
          subtitle = "First run") +
  geom_point(aes(colour = factor(SampID))) +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")+
  guides(colour="none", size="none")

ggplotly(p_gmo)
```


