---
title: "Splitting Data for NiN v 3 analyses - R Notebook"
output: html_notebook
---

#INPUTS
Assumes you hve already run at least the scripts:
- workflow_deepseaNiN_Start
- Workflow_deepseaNiN_Start_splitBiogeog

(Else further subdivisions of the same.) 

Requires the main analysis environment to be loaded.

# Selecting geographic subsets
This is to explore thresholds/relationships locally. E.g. To look for threshold values/ local relationships before deciding on splits. You need to already have a theory that you want to test in order to select the geographic regions to look at.


## OPTIONAL: load new dataset and apply splits SKIP IF ALREADY HAVE "joinedDat"


```{r}
#NB 2022-09-29 was the topcopy (all sample/Oslo) dataset being used before the latest species cleaning
#2022-10-19 latest (all sample/Oslo) species cleaning applied
# USE DIFFERENT SCRIPT FOR 1perR station dataset

env_orig <- read.csv(file.path(dataPath,"inputs/envDataOslo_2022-10-19.csv")) %>% as.data.frame
otu_orig <- read.csv(file.path(dataPath,"inputs/widedataOslo_2022-10-19.csv")) %>% as.data.frame



#left join and split to ensure ordered identically

#joinedDat<-otu_orig%>% dplyr::left_join(.,env_orig, by="SampID")
joinedDat<- left_join(otu_orig,env_orig, by=c("SampID"="SampID2"))

```




## MAIN SPLITS

### Density threshold splits

density split 1031.8 threshold

```{r}
joinedDat$swDensRob_avs<-swRho(salinity=joinedDat$Smean_Robinson,
                         temperature=joinedDat$Tmean_Robinson,
                         pressure=(joinedDat$bathy*-1),
                         eos="unesco")

loDensShallowJoinedDat<-subset(joinedDat,swDensRob_avs<1031.1)
plot(loDensShallowJoinedDat$X.y, loDensShallowJoinedDat$Y)
LDensShlw_otu_sort<-loDensShallowJoinedDat %>% select(c(2:391))
LDensShlw_env_sort<-loDensShallowJoinedDat %>% select(c(2,397:532))                                


hiDensDeepJoinedDat<-subset(joinedDat,swDensRob_avs>1031.1)
plot(hiDensDeepJoinedDat$X.y, hiDensDeepJoinedDat$Y)
HDensDeep_otu_sort<-hiDensDeepJoinedDat %>% select(c(2:391))
HDensDeep_env_sort<-hiDensDeepJoinedDat %>% select(c(2,397:532))  
```


##### SAVE DATASETS
```{r}
write.csv(HDensDeep_otu_sort, file=file.path(dataPath,(paste0("inputs/HDensDeep_otu_sort_", Sys.Date(),".csv"))))
write.csv(HDensDeep_env_sort, file=file.path(dataPath,(paste0("inputs/HDensDeep_env_sort_", Sys.Date(),".csv"))))

write.csv(LDensShlw_otu_sort, file=file.path(dataPath,(paste0("inputs/LDensShlw_otu_sort_", Sys.Date(),".csv"))))
write.csv(LDensShlw_env_sort, file=file.path(dataPath,(paste0("inputs/LDensShlw_env_sort_", Sys.Date(),".csv"))))
```


### Split HiDens at 1500m

```{r}

hiDensDeepJoinedDat$bathyPositive<-(hiDensDeepJoinedDat$bathy*-1)

DeepBelow1500<-subset(hiDensDeepJoinedDat,bathyPositive>1500)
DeepAbove1500<-subset(hiDensDeepJoinedDat,bathyPositive<1500)
DeepBelow1500<-DeepBelow1500%>%select(-c("bathyPositive"))
DeepAbove1500<-DeepAbove1500%>%select(-c("bathyPositive"))

plot(DeepBelow1500$X.y, DeepBelow1500$Y)
DeepBelow1500_otu_sort<-DeepBelow1500 %>% select(c(2:391))
DeepBelow1500_env_sort<-DeepBelow1500 %>% select(c(2,397:532))  


plot(DeepAbove1500$X.y, DeepAbove1500$Y)
DeepAbove1500_otu_sort<-DeepAbove1500 %>% select(c(2:391))
DeepAbove1500_env_sort<-DeepAbove1500 %>% select(c(2,397:532))  
```

##### SAVE DATASETS
```{r}
write.csv(DeepBelow1500_otu_sort, file=file.path(dataPath,(paste0("inputs/DeepBelow1500_otu_sort_", Sys.Date(),".csv"))))
write.csv(DeepBelow1500_env_sort, file=file.path(dataPath,(paste0("inputs/DeepBelow1500_env_sort_", Sys.Date(),".csv"))))

write.csv(DeepAbove1500_otu_sort, file=file.path(dataPath,(paste0("inputs/DeepAbove1500_otu_sort_", Sys.Date(),".csv"))))
write.csv(DeepAbove1500_env_sort, file=file.path(dataPath,(paste0("inputs/DeepAbove1500_env_sort_", Sys.Date(),".csv"))))
```



### REMOVE SPITSBERGENBANKEN OUTLIERS FROM LowDens Dataset

c("1805_01","1801_00","1802_02", "1799_01", "1803_00", "1798_02","1800_00","1995_00")

```{r}
#SB<-c("1805_01","1801_00","1802_02", "1799_01", "1803_00", "1798_02","1800_00","1995_00")

plot(loDensShallowJoinedDat$X.y, loDensShallowJoinedDat$Y)

#loDensNoSBJoinedDat <-loDensShallowJoinedDat %>% filter(!(SampID %in% c("1805_01",
                                                                        # "1801_00",
                                                                        # "1802_02", 
                                                                        # "1799_01", 
                                                                        # "1803_00", 
                                                                        # "1798_02",
                                                                        # "1800_00",
                                                                        # "1995_00")))
loDensShallowJoinedDat$bathyPositive<-(loDensShallowJoinedDat$bathy*-1)
loDensNoSBJoinedDat <-subset(loDensShallowJoinedDat, bathyPositive>100)
loDensNoSBJoinedDat<-loDensNoSBJoinedDat%>%select(-c("bathyPositive"))



plot(loDensNoSBJoinedDat$X.y, loDensNoSBJoinedDat$Y)


LDensNoSB_otu_sort<-loDensNoSBJoinedDat %>% select(c(2:391))
LDensNoSB_env_sort<-loDensNoSBJoinedDat %>% select(c(2,397:532))  

```
##### SAVE DATASETS
```{r}
write.csv(LDensNoSB_otu_sort, file=file.path(dataPath,(paste0("inputs/LDensNoSB_otu_sort_", Sys.Date(),".csv"))))
write.csv(LDensNoSB_env_sort, file=file.path(dataPath,(paste0("inputs/LDensNoSB_env_sort_", Sys.Date(),".csv"))))
```


### Split LoDensNoSB by dissox

```{r}
#SB<-c("1805_01","1801_00","1802_02", "1799_01", "1803_00", "1798_02","1800_00","1995_00")

plot(loDensNoSBJoinedDat$X.y, loDensNoSBJoinedDat$Y)


loDensNoSBhiO2JoinedDat<-subset(loDensNoSBJoinedDat,BO22_dissoxmean_bdmean>305)
plot(loDensNoSBhiO2JoinedDat$X.y, loDensNoSBhiO2JoinedDat$Y)
LDnoSBhi02_otu_sort<-loDensNoSBhiO2JoinedDat %>% select(c(2:391))
LDnoSBhi02_env_sort<-loDensNoSBhiO2JoinedDat %>% select(c(2,397:532))  


loDensNoSBloO2JoinedDat<-subset(loDensNoSBJoinedDat,BO22_dissoxmean_bdmean<305)
plot(loDensNoSBloO2JoinedDat$X.y, loDensNoSBloO2JoinedDat$Y)
LDnoSBlo02_otu_sort<-loDensNoSBloO2JoinedDat %>% select(c(2:391))
LDnoSBlo02_env_sort<-loDensNoSBloO2JoinedDat %>% select(c(2,397:532))    

```
##### SAVE DATASETS
```{r}
write.csv(LDnoSBlo02_otu_sort, file=file.path(dataPath,(paste0("inputs/LDnoSBlo02_otu_sort_", Sys.Date(),".csv"))))
write.csv(LDnoSBlo02_env_sort, file=file.path(dataPath,(paste0("inputs/LDnoSBlo02_env_sort_", Sys.Date(),".csv"))))

write.csv(LDnoSBhi02_otu_sort, file=file.path(dataPath,(paste0("inputs/LDnoSBhi02_otu_sort_", Sys.Date(),".csv"))))
write.csv(LDnoSBhi02_env_sort, file=file.path(dataPath,(paste0("inputs/LDnoSBhi02_env_sort_", Sys.Date(),".csv"))))
```



--------------

## OPTIONAL: eplore an interactive plot to aid split choices

### Choose geographic areas:
Choosing the exact areas can be done using the interactive "ggplotly" XY map of temperature to zoom in on the area concerned and note the bounding box values to use.

That is reproduced here:
##### X v Y - temp
```{r}
tb_ax1<- ggplot(data = env_sub_meta,
                      aes(x = X,
                          y = Y)) +
  theme_classic() +
  geom_point(aes(colour = Tmean_Robinson),
             size = 1) +
  scale_colour_gradientn(limits = c(-2, 10),
                        colors=c('blue','green','red'))+
  ggtitle("Geography (X v Y) - coloured by mean temp")

ggplotly(tb_ax1)
```

This gives us the following bounding box selections

NorlandSlope: X:205261 - 364461, Y: 743294 - 7621734
FinnmarkCoast: X: 599461 - 1138061, Y: 7794534 - 8013534
MareanoØst: X: 1009861 - 1162261, Y: 7834134 - 8505934

Additional selections made after investigating these three areas were made to check the temperature thrshold on the slope in the north and south:

SvalbardSlope: X: 361061 - 561862, Y: 7915734 - 8782734
MøreSlope: X: -47738 - 93861, Y: 6976734 - 7234534

After these explorations it was decided to generate a subset by density. A threshold of 1031.8 was used to split the dataset.

## Make data subsets to explore

### Geographic subsets to explore temp gradients

```{r}
NorlandSlopeJoinedDat<-subset(joinedDat, X.y>205261 & X.y<364461 & Y>743294 & Y<7621734)
plot(NorlandSlopeJoinedDat$X.y, NorlandSlopeJoinedDat$Y)
NSotu_sort<-NorlandSlopeJoinedDat %>% select(c(2:391))
NSenv_sort<-NorlandSlopeJoinedDat %>% select(c(2,397:532))

FinnmarkCoastJoinedDat<-subset(joinedDat, X.y>599461 & X.y<1138061 & Y>7794534 & Y<8013534)
plot(FinnmarkCoastJoinedDat$X.y, FinnmarkCoastJoinedDat$Y)
FCotu_sort<-FinnmarkCoastJoinedDat %>% select(c(2:391))
FCenv_sort<-FinnmarkCoastJoinedDat %>% select(c(2,397:532))

MareanoØstJoinedDat<-subset(joinedDat, X.y>1009861 & X.y<1162261 & Y>7834134 & Y<8505934)
plot(MareanoØstJoinedDat$X.y, MareanoØstJoinedDat$Y)
MØotu_sort<-MareanoØstJoinedDat %>% select(c(2:391))
MØenv_sort<-MareanoØstJoinedDat %>% select(c(2,397:532))


```

##### SAVE DATASETS
```{r}
write.csv(NSotu_sort, file=file.path(dataPath,(paste0("inputs/NSotu_sort_", Sys.Date(),".csv"))))
write.csv(NSenv_sort, file=file.path(dataPath,(paste0("inputs/NSenv_sort_", Sys.Date(),".csv"))))

write.csv(FCotu_sort, file=file.path(dataPath,(paste0("inputs/FCotu_sort_", Sys.Date(),".csv"))))
write.csv(FCenv_sort, file=file.path(dataPath,(paste0("inputs/FCenv_sort_", Sys.Date(),".csv"))))

write.csv(MØotu_sort, file=file.path(dataPath,(paste0("inputs/MOotu_sort_", Sys.Date(),".csv"))))
write.csv(MØenv_sort, file=file.path(dataPath,(paste0("inputs/MOenv_sort_", Sys.Date(),".csv"))))

```


### Selections to explore deep temp splits
```{r}
SvalbardSlopeJoinedDat<-subset(joinedDat, X.y>361061 & X.y<561862 & Y>7915734 & Y<8782734)
plot(SvalbardSlopeJoinedDat$X.y, SvalbardSlopeJoinedDat$Y)
SSotu_sort<-SvalbardSlopeJoinedDat %>% select(c(2:391))
SSenv_sort<-SvalbardSlopeJoinedDat %>% select(c(2,397:532))

MøreSlopeJoinedDat<-subset(joinedDat, X.y>-47738 & X.y<93861 & Y>6976734 & Y<7234534)
plot(MøreSlopeJoinedDat$X.y, MøreSlopeJoinedDat$Y)
MSotu_sort<-MøreSlopeJoinedDat %>% select(c(2:391))
MSenv_sort<-MøreSlopeJoinedDat %>% select(c(2,397:532))
```

##### SAVE DATASETS
```{r}
write.csv(SSotu_sort, file=file.path(dataPath,(paste0("inputs/SSotu_sort_", Sys.Date(),".csv"))))
write.csv(SSenv_sort, file=file.path(dataPath,(paste0("inputs/SSenv_sort_", Sys.Date(),".csv"))))

write.csv(MSotu_sort, file=file.path(dataPath,(paste0("inputs/MSotu_sort_", Sys.Date(),".csv"))))
write.csv(MSenv_sort, file=file.path(dataPath,(paste0("inputs/MSenv_sort_", Sys.Date(),".csv"))))
```

