---
title: "MAREANO - NiN: First run"
authors: Thijs van Son, Rune Halvorsen, Rebecca Ross, Genoveva Gonzales-Mirelis, Margaret Dolan
date: "Last Rendered on `r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 2
    toc_float: yes
    fig_width: 7
    fig_height: 7
always_allow_html: true 
---

## Packages
```{r}
library(dplyr)
library(ggcorrplot)
library(ggpubr)
library(patchwork)
# library("here")
# library("bookdown")
# library("downloadthis")
library(vegan)
library(plyr)
library(e1071)
library(tidyverse)
library(viridis)
library(GGally)
library(ggrepel)
library(ggordiplots)
library(readr)
library(RColorBrewer)


source("gg_ordisurf_viridis.R")
```


## Data
```{r}

env_orig <- read.csv(file.path(dataPath,"envData_2022-09-13.csv")) %>% as.data.frame
otu_orig <- read.csv(file.path(dataPath,"widedata_2022-09-13.csv")) %>% as.data.frame

#left join and split to ensure ordered identically

#joinedDat<-otu_orig%>% dplyr::left_join(.,env_orig, by="SampID")
joinedDat<- left_join(otu_orig,env_orig, by=c("SampID"="SampID2"))

otu_sort<-joinedDat %>% select(c(2:351))
env_sort<-joinedDat %>% select(c(2,352:491))

# Sorting and ordering by SampID
# env_orig <- env_orig[order(env_orig$SampID), ]
# otu_orig <- otu_orig[order(otu_orig$SampID), ]

table(env_sort$SampID2 == otu_sort$SampID)


#must be same length and equal to unique sample number length - the ordered lists are assumed to be directly relatable on a row by row basis

dim(env_sort)
dim(otu_sort)
```


### Data cleaning
```{r}
## Removing NAs
otuCompl <- otu_sort[complete.cases(env_sort[, -c(1:20)]), ]
envCompl <- env_sort[complete.cases(env_sort[, -c(1:20)]), ]

## Removing observations with less than 4 OTUs
sel <- rowSums(otuCompl[, -c(1:2)]) >= 4
otuSel <- otuCompl[sel, ]
envSel <- envCompl[sel, ]

## Removing phosphate
#envSel <- envSel %>% select(-phosphate_mean.tif)

dim(otuSel); dim(envSel)
```


### [Optional] Data thinning
```{r, eval=FALSE}
# otu_red <- otu1[1:500, ]
# otu <- otu_red
# 
# env_red <- env[1:500, ]
# env <- env_red
```

### Whole dataset for first run
```{r}
otu<-otuSel[,-c(1)]
env<-envSel[, -c(1:2,4:20)]

table(is.na(otu))

table(is.na(env))

#str(otuSel)
#str(envSel)
```


### Splitting in subsets
```{r}
# ## Class 1
# otu1 <- subset(otuSel, envSel$SplitRev == 1)
# env1 <- envSel %>% filter(SplitRev == 1)
# 
# ## Class 2
# otu2 <- subset(otuSel, envSel$SplitRev == 2)
# env2 <- envSel %>% filter(SplitRev == 2)
# 
# ## Class 4
# otu3 <- subset(otuSel, envSel$SplitRev == 3)
# env3 <- envSel %>% filter(SplitRev == 3)
# 
# ## Class 4
# otu4 <- subset(otuSel, envSel$SplitRev == 4)
# env4 <- envSel %>% filter(SplitRev == 4)
# 
# ## Class 6
# otu6 <- subset(otuSel, envSel$SplitRev == 6)
# env6 <- envSel %>% filter(SplitRev == 6)
# 
# ## Class 7
# otu7 <- subset(otuSel, envSel$SplitRev == 7)
# env7 <- envSel %>% filter(SplitRev == 7)
# 
# ## Class 8
# otu8 <- subset(otuSel, envSel$SplitRev == 8)
# env8 <- envSel %>% filter(SplitRev == 8)
```

## Selecting subset
```{r}
# ## Selecting 
# otu <- otu2
# env <- env2

```

## format data correctly
```{r}
#env$coords.x1<-as.numeric(env$coords.x1)
#env$coords.x2<-as.numeric(env$coords.x2)


```


## Abundance weighting
Make function
You might have to drop variables that have been imported as character
```{r}

otu_pa <- decostand(x = otu[, -c(1)],
                    method = "pa")

# y = ax^w         # power transformation formula
dt <- otu[, -c(1:2)]          # species data to transform
x_mn <- min(dt[dt > 0])
x_mx <- max(dt)
rng <- 6           # abundance range
w <- log(rng) / (log(x_mx) - log(x_mn))
a <- x_mn^(-w)
# otu_6 <- a * dt[, -c(1:3)]^w
otu_6 <- a * dt^w
range(otu_6)
```

## Ordination methods
### DCA PA
```{r}
dca_pa <- decorana(veg = otu_pa)
```


### GNMDS PA
#### Distances
```{r, eval=FALSE}
## Bray-Curtis
dist_pa <- vegdist(x = otu_pa, method = "bray")

## Geodist
ep <- 0.8     # epsilon
geodist_pa <- isomapdist(dist = dist_pa, epsilon = ep)

saveRDS(geodist_pa,
        file = (file.path(dataPath,"geodist_pa.rds")))
```

#### Ordination
```{r, eval=FALSE}
# monoMDS
d <- 2
mds_pa <- list()

for (i in 1:100) {
  mds_pa[[i]]<-monoMDS(geodist_pa,
                    matrix(c(runif(dim(otu_pa)[1]*d)),
                           nrow = dim(otu_pa)[1]),
                    k = d,
                    model = "global",
                    maxit = 2000,
                    smin = 1e-7,
                    sfgrmin = 1e-7)
}

saveRDS(mds_pa,
        file = (file.path(dataPath,"mds_pa.rds")
```

##### Best nmds solution - PA
Make function?
```{r, eval=TRUE}
# Loading geodist object
# geodist_nfi <- readRDS(file = "../SplitRev2_geodist_pa_full.rds")

# Loading mds results
# mds <- readRDS(file = "../SplitRev2_mds_pa_full.rds")

## Extracting the stress of each nmds iteration
mds_stress_pa<-unlist(lapply(mds_pa, function(v){v[[22]]})) 

ordered_pa <-order(mds_stress_pa)

## Best, second best, and worst solution
mds_stress_pa[ordered_pa[1]]
mds_stress_pa[ordered_pa[2]]
mds_stress_pa[ordered_pa[10]]

## Scaling of axes to half change units and varimax rotation by postMDS
mds_best_pa<-postMDS(mds_pa[[ordered_pa[1]]],
                  geodist_pa, 
                  pc = TRUE, 
                  halfchange = TRUE, 
                  threshold = ep)     # Is this threshold related to the epsilon above?
mds_best_pa

mds_secbest_pa <- postMDS(mds_pa[[ordered_pa[2]]],
                          geodist_pa, 
                          pc = TRUE, 
                          halfchange = TRUE, 
                          threshold = ep)
mds_secbest_pa

## Procrustes comparisons
procr_pa <- procrustes(mds_best_pa,
                    mds_secbest_pa,
                    permutations=999)
protest(mds_best_pa,
        mds_secbest_pa,
        permutations=999)

plot(procr_pa)

png(file=file.path(dataPath,"procrustes_pa.png"), width=1000, height=700)
plot(procr_pa)
dev.off()
```

##### Correlation of axis: DCA vs NMDS - PA
```{r}
# Extracting ordination axis
ax <- 2
axis_pa <- cbind(mds_best_pa$points,
                 scores(dca_pa,
                        display = "sites",
                        origin = TRUE)[, 1:ax])

ggcorr(axis_pa, 
       method=c("everything","kendall"), 
       label = TRUE,
       label_size = 3, 
       label_color = "black",  
       nbreaks = 8,
       label_round = 3,
       low = "red",
       mid = "white",
       high = "green")

ggsave(filename = file.path(dataPath,"correlationPCAvsNMDS_PA.png"),
       device = "png",
       dpi=300 )

# Switching direction of NMDS1
# mds_best$points[, 1] <- -mds_best$points[, 1]

```


### DCA R6
```{r}
dca_r6 <- decorana(veg = otu_6)
```


### GNMDS R6
#### Distances
```{r}
## Bray-Curtis
dist_r6 <- vegdist(x = otu_6, method = "bray")

## Geodist
ep <- 0.80     # epsilon
geodist_r6 <- isomapdist(dist = dist_r6, epsilon = ep)
```

#### Ordination
```{r, eval=FALSE}
# monoMDS
d <- 2
mds_r6 <- list()

for (i in 1:200) {
  mds_r6[[i]]<-monoMDS(geodist_r6,
                    matrix(c(runif(dim(otu_6)[1]*d)),
                           nrow = dim(otu_6)[1]),
                    k = d,
                    model = "global",
                    maxit = 2000,
                    smin = 1e-7,
                    sfgrmin = 1e-7)
}

saveRDS(mds_pa,
        file = file.path(dataPath,"mds_r6.rds")
```

##### Best nmds solution - PA
```{r, eval=TRUE}
# Loading geodist object
# geodist_nfi <- readRDS(file = "../SplitRev2_geodist_nfi.rds")

# Loading mds results
# mds <- readRDS(file = "../SplitRev2_mds.rds")

## Extracting the stress of each nmds iteration
mds_stress_r6<-unlist(lapply(mds_r6, function(v){v[[22]]})) 

ordered_r6 <-order(mds_stress_r6)

## Best, second best, and worst solution
mds_stress_r6[ordered_r6[1]]
mds_stress_r6[ordered_r6[2]]
mds_stress_r6[ordered_r6[100]]

## Scaling of axes to half change units and varimax rotation by postMDS
mds_best_r6<-postMDS(mds_r6[[ordered_r6[1]]],
                     geodist_r6, 
                     pc = TRUE, 
                     halfchange = TRUE, 
                     threshold = ep)     # Is this threshold related to the epsilon above?
mds_best_r6

mds_secbest_r6<-postMDS(mds_r6[[ordered_r6[2]]],
                        geodist_r6, 
                        pc = TRUE, 
                        halfchange = TRUE, 
                        threshold = ep)
mds_secbest_r6

## Procrustes comparisons
procr_r6 <- procrustes(mds_best_r6,
                       mds_secbest_r6,
                       permutations=999)
protest(mds_best_r6,
        mds_secbest_r6,
        permutations=999)

plot(procr_r6)

png("SplitRev2_procrustes_r6.png", width=1000, height=700,)
plot(procr_r6)
dev.off()
```

##### Correlation of axis: DCA vs NMDS - PA
```{r}
# Extracting ordination axis
ax <- 2
axis_r6 <- cbind(mds_best_r6$points,
                 scores(dca_r6,
                        display = "sites",
                        origin = TRUE)[, 1:ax])

ggcorr(axis_r6, 
       method=c("everything","kendall"), 
       label = TRUE,
       label_size = 3, 
       label_color = "black",  
       nbreaks = 8,
       label_round = 3,
       low = "red",
       mid = "white",
       high = "green")

ggsave(filename = "SplitRev2_correlationPCAvsNMDS_r6.png",
       device = "png",
       dpi=300 )

# Switching direction of NMDS1
# mds_best$points[, 1] <- -mds_best$points[, 1]

```

### Plotting DCA & GNMDS
```{r}
## Adding scores to data frame
otu_6$gnmds1 <- mds_best_r6$points[, 1]
otu_6$gnmds2 <- mds_best_r6$points[, 2]
otu_6$dca1 <- scores(dca_r6, display = "sites", origin = TRUE)[, 1]
otu_6$dca2 <- scores(dca_r6, display = "sites", origin = TRUE)[, 2]

p_gnmds_r6 <- ggplot(data = otu_6,
                     aes(x = gnmds1,
                         y = gnmds2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("GNMDS",
          subtitle = "First run") +
  geom_point(colour = "red") +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")

p_dca_r6 <- ggplot(data = otu_6,
                   aes(x = dca1,
                       y = -dca2)) +
  theme_classic() +
  coord_fixed() +
  ggtitle("DCA",
          subtitle = "First run") +
  geom_point(colour = "red") +
  geom_vline(xintercept = 0,
             linetype = 2,
             colour = "lightgray") +
  geom_hline(yintercept = 0,
             linetype = 2,
             colour = "lightgray")

p_gnmds_r6 + p_dca_r6

ggsave("SplitRev2_gnmds_dca.png",
       device = "png", 
       dpi=300)

```

## Species-environment relationships
### Selecting ordination
```{r}
ord <- mds_best_r6

## Axis scores if selected ord is GNMDS
axis <- ord$points %>% as.data.frame

## Axis scores if selected ord is DCA
# axis <- scores(ord,
#                display = "sites",
#                origin = TRUE)[, 1:ax])

```

### Correlation ordination axes and environmental variables
#### Removing non-env vars
```{r}
env_corr <- env %>% select(-c(Lnr,
                              Tot1810,
                              SplitRaw,
                              Split,
                              SplitRev,
                              RegionSimp,
                              SampID,
                              samp,
                              no,
                              T,
                              B,
                              class))

env_corr$coords.x1<-as.numeric(env_corr$coords.x1)
env_corr$coords.x2<-as.numeric(env_corr$coords.x2)

str(env_corr)
```

#### Correlations
```{r}
# Vector to hold correlations
cor_ax1 <- NULL
cor_ax2 <- NULL
pv_ax1 <- NULL
pv_ax2 <- NULL

# NMDS1
for( i in seq(length(env_corr))) {
  ct.i <- cor.test(axis$MDS1,
                   env_corr[, i],
                   method = "kendall")
  cor_ax1[i] <- ct.i$estimate
  pv_ax1[i] <- ct.i$p.value
}

# NMDS2
for( i in seq(length(env_corr))) {
  ct.i <- cor.test(axis$MDS2,
                   env_corr[, i],
                   method = "kendall")
  cor_ax2[i] <- ct.i$estimate
  pv_ax2[i] <- ct.i$p.value
}

cor_tab <- data.frame(env = names(env_corr),
                      ord_ax1 = cor_ax1,
                      pval_ax1 = pv_ax1,
                      ord_ax2 = cor_ax2,
                      pval_ax2 = pv_ax2)

cor_tab

write.csv(x = cor_tab,
          file = "SplitRev2_cor-table_r6_200rep.csv",
          row.names = FALSE)
```

### Sel env var & removal NAs
```{r}
env_os <- env[, c("bathy.tif",
                  "hard.tif",
                  "dissox_mean.tif",
                  "mud.tif",
                  "salt_min.tif",
                  "salt_std.tif",
                  "coords.x2",
                  "spd_mean.tif",
                  "gelatinous_snow",
                  "temp_mean.tif",
                  "pebble",
                  "icecover_mean.tif",
                  "gravel.tif")]
env_os
str(env_os)

```


### Ordisurfs
```{r}
ordsrfs <- list(length = ncol(env_os))

for (i in seq(ncol(env_os))) {
  os.i <- gg_ordisurf(ord = ord,
                      env.var = env_os[, i],
                      pt.size = 1,
                      # binwidth = 0.05,
                      var.label = names(env_os)[i],
                      gen.text.size = 10,
                      title.text.size = 15,
                      leg.text.size = 10)
  
  ordsrfs[[i]] <- os.i$plot
}

ordsrfs_plt <- ggarrange(plotlist = ordsrfs,
                         nrow = 5,
                         ncol = 3)

ordsrfs_plt

ggexport(ordsrfs_plt,
          filename = "SplitRev2_ordisurfs_full_sel.png",
          width = 1000,
          height = 2000)

png("SplitRev2_ordisurfs_full_sel.png", width=2000, height=2200)
ordsrfs_plt
dev.off()
```

### Envfit
```{r}
## Envfot plot
gg_envfit(ord = ord,
          env = env_os,
          pt.size = 1)

ggsave(filename = "SplitRev2_EnvFit.png",
       device = "png",
       dpi=300 )


## Envfit analysis

ef <- envfit(ord = ord,
             env = env_os,
             # na.rm = TRUE
             )
efDF <- as.data.frame(scores(ef,
                             display = "vectors"))
```


### Data split plot
#### Attaching ordination scores
```{r}
env$gnmds1 <- axis$MDS1
env$gnmds2 <- axis$MDS2
```

#### Plots
```{r}
TB_splitRaw <- ggplot(data = env,
                      aes(x = temp_mean.tif,
                          y = bathy.tif)) +
  theme_classic() +
  geom_point(aes(colour = SplitRaw),
             size = 2) +
  scale_colour_brewer(palette = "Paired") +
  # geom_vline(xintercept = 0) +
  # geom_segment(aes(x = 2, y = 0,
  #                  xend = 2, yend = -675)) +
  # geom_segment(aes(x = 0, y = -700,
  #                  xend = 5, yend = -575)) +
  # geom_segment(aes(x = 0, y = -700,
  #                  xend = -1.5, yend = -700)) +
  # geom_segment(aes(x = 0, y = -425,
  #                  xend = 2, yend = -325)) +
  # geom_segment(aes(x = 2, y = -500,
  #                  xend = 5, yend = -400)) +
  # geom_segment(aes(x = 5, y = -400,
  #                  xend = 8, yend = 0)) +
  # geom_segment(aes(x = 0, y = -425,
  #                  xend = -1.5, yend = -425)) +
  # annotate(geom = "text", x = -1, y = -1500, label = "1", size = 7) +
  # annotate(geom = "text", x = -1, y = -75, label = "2", size = 7) +
  # annotate(geom = "text", x = 2.5, y = -1000, label = "3", size = 7) +
  # annotate(geom = "text", x = 1, y = -550, label = "4", size = 7) +
  # annotate(geom = "text", x = 0.5, y = -250, label = "5", size = 7) +
  # annotate(geom = "text", x = 6, y = -500, label = "6", size = 7) +
  # annotate(geom = "text", x = 4.5, y = -250, label = "7", size = 7) +
  # annotate(geom = "text", x = 1.1, y = 50, label = "8", size = 7) +
  ggtitle("Temp (mean) vs Bathy - coloured by SplitRaw")

TB_split <- ggplot(data = env,
                      aes(x = temp_mean.tif,
                          y = bathy.tif)) +
  theme_classic() +
  geom_point(aes(colour = Split),
             size = 2) +
  scale_colour_brewer(palette = "Paired") +
  # geom_vline(xintercept = 0) +
  # geom_segment(aes(x = 2, y = 0,
  #                  xend = 2, yend = -675)) +
  # geom_segment(aes(x = 0, y = -700,
  #                  xend = 5, yend = -575)) +
  # geom_segment(aes(x = 0, y = -700,
  #                  xend = -1.5, yend = -700)) +
  # geom_segment(aes(x = 0, y = -425,
  #                  xend = 2, yend = -325)) +
  # geom_segment(aes(x = 2, y = -500,
  #                  xend = 5, yend = -400)) +
  # geom_segment(aes(x = 5, y = -400,
  #                  xend = 8, yend = 0)) +
  # geom_segment(aes(x = 0, y = -425,
  #                  xend = -1.5, yend = -425)) +
  # annotate(geom = "text", x = -1, y = -1500, label = "1", size = 7) +
  # annotate(geom = "text", x = -1, y = -75, label = "2", size = 7) +
  # annotate(geom = "text", x = 2.5, y = -1000, label = "3", size = 7) +
  # annotate(geom = "text", x = 1, y = -550, label = "4", size = 7) +
  # annotate(geom = "text", x = 0.5, y = -250, label = "5", size = 7) +
  # annotate(geom = "text", x = 6, y = -500, label = "6", size = 7) +
  # annotate(geom = "text", x = 4.5, y = -250, label = "7", size = 7) +
  # annotate(geom = "text", x = 1.1, y = 50, label = "8", size = 7) +
  ggtitle("Temp (mean) vs Bathy - coloured by Split")

EN_split <- ggplot(data = env,
              aes(x = coords.x1,
                  y = coords.x2)) +
  theme_classic() +
  geom_point(aes(colour = Split),
             size = 2) +
  scale_colour_brewer(palette = "Paired") +
  ggtitle("Easting vs Northing - coloured by Split")

EN_split + TB_split



```

### Coordinates coloured by GNMDS1 & 2
```{r}
EN1 <- ggplot(data = env,
              aes(x = coords.x1,
                  y = coords.x2)) +
  theme_classic() +
  geom_point(aes(colour = gnmds1),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Easting vs Northing - coloured by GNMDS1")

EN2 <- ggplot(data = env,
              aes(x = coords.x1,
                  y = coords.x2)) +
  theme_classic() +
  geom_point(aes(colour = gnmds2),
             size = 2) +
  scale_colour_gradient2(low = "red",
                         mid = "yellow",
                         high = "green") +
  ggtitle("Easting vs Northing - coloured by GNMDS2")

EN_temp <- ggplot(data = env,
              aes(x = coords.x1,
                  y = coords.x2)) +
  theme_classic() +
  geom_point(aes(colour = temp_mean.tif),
             size = 2) +
  scale_colour_gradient2(low = "Blue",
                         mid = "yellow",
                         high = "red",
                         midpoint = 4) +
  ggtitle("Easting vs Northing - coloured by temp")

EN1 + EN2
```


## Convert to r script
```{r}
library(knitr)
knitr::purl(input = "first-run.Rmd",
            output = "first-run.r")
```

